### 2021-12-12 draft_yma.Rmd
data exploration, distribution, outlier, correlation

---
title: "draft_yma"
author: "Anna Ma"
date: "12/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
```

## Data exploration

 **Q: do we need to convert other variables to per 1,000 population? like docs, beds, etc.**

```{r, message=FALSE}
cdi = read_csv("cdi.csv") %>%
  janitor::clean_names() %>%
  mutate(crm_1000 = 1000*(crimes/pop),
         # convert region to factors and recoded them accordingly 
         region = factor(region, levels = 1:4,
                    labels = c("northeast", "north central", "south", "west"))) %>% view()

```

### Descriptive statistics of all varaibles
```{r statistic summary}

cdi_descriptive = cdi %>% select(-c(id,cty,state,region))

# Global
skimr::skim(cdi_descriptive) %>% 
  select(-c("skim_type","complete_rate")) %>% 
  knitr::kable(
    col.names = c("variable", "n_missing", "mean","sd","min","Q25","median","Q75","max","histogram"),
    caption = "Global Summary", digits = 2)
```

Q: 1. do we need to group them by state or county? group by gounty gives 5000+ rows though...
  2. do we need box plot still? 
  3. again, transfer some variables to "per pop" / "per 1000 pop" ?

### Descriptive analysis

We can use the box plot/ or histogram to check for normality. But I forgot when do we need normality...isn't it for residual??

```{r}
par(mfrow=c(3,5))

boxplot(cdi$area,  main = "area")
boxplot(cdi$pop, main = "pop")
boxplot(cdi$pop18, main = "pop18")
boxplot(cdi$pop65, main = "pop65")

boxplot(cdi$docs, main = "docs")
boxplot(cdi$beds, main = "beds")
boxplot(cdi$crimes, main = "crimes")
boxplot(cdi$crm_1000,  main = "crm_1000")

boxplot(cdi$hsgrad, main = "hsgrad")
boxplot(cdi$bagrad, main = "bagrad")
boxplot(cdi$poverty, main = "poverty")
boxplot(cdi$unemp,main = "unemp")
boxplot(cdi$pcincome, main = "pcincome")
boxplot(cdi$totalinc, main = "totalinc")

```

### Correlation

#### Pairwise relationship

* This gives us an idea of the correlation between each variable, but my old project build whole model first then assessed the correlation. Need Discussion!

```{r}
library(corrplot)
corrplot(cor(cdi_descriptive), type = "upper", diag = FALSE)
```

* Here is the correlation with pairs function. More specific than the heat map above. We can observe the correlation bewteen all the terms here
```{r}
pairs(crm_1000 ~.,data=cdi_descriptive, panel = panel.smooth, upper.panel = NULL, main = "Scatterplot Matrix")
```

#### Marginal distribution ?

### Distribution of outcome

```{r}
cdi %>% 
  ggplot(aes(x = crm_1000)) +
  geom_histogram()

# log transfor the outcome
cdi %>% 
  ggplot(aes(x = log(crm_1000))) +
  geom_histogram()

```

do we look at the distribution of outcome like this and transform them here? check again

### States with unusual rates

group by state
```{r}
cdi_by_state = cdi %>% group_by(state) %>% 
  summarise(state_pop = sum(pop),
            state_crimes = sum(crimes)) %>% 
  mutate(state_CRM_1000 = 1000*state_crimes/state_pop)
```

Not sure what counts as out lier, this is what the other group did.

```{r}
upper = quantile(cdi$crm_1000, 0.75)
lower = quantile(cdi$crm_1000, 0.25)
IQR = upper - lower
cdi %>% 
  filter(crm_1000 > upper + 1.5*IQR,
         crm_1000 > lower - 1.5*IQR) %>% 
  dplyr::select(cty, crm_1000) %>%
  knitr::kable(digits = 2)
```

This shows no outlier on the state level tho...

```{r}
upper = quantile(cdi_by_state$state_CRM_1000, 0.75)
lower = quantile(cdi_by_state$state_CRM_1000, 0.25)
IQR = upper - lower
cdi_by_state %>% 
  filter(state_CRM_1000 > upper + 1.5*IQR,
         state_CRM_1000 > lower - 1.5*IQR) %>% 
  dplyr::select(state, state_CRM_1000) %>%
  knitr::kable(digits = 2)
boxplot(cdi_by_state$state_CRM_1000, main = 'State Crime Rate per 1000 people')
```

### 2021-12-13 draft_yma.Rmd
marginal distribution

@@
 
- **Q: do we need to convert other variables to per 1,000 population? like docs, beds, etc.**
+ **converted variables to per person and docs, beds, etc.**
 
@@
   mutate(crm_1000 = 1000*(crimes/pop),
+         pdocs = docs/pop,
+         pbeds = beds/pop,
+         density_pop = pop/area,
          # convert region to factors and recoded them accordingly 
          region = factor(region, levels = 1:4,
-                    labels = c("northeast", "north central", "south", "west"))) %>% view()
+                    labels = c("northeast", "north central", "south", "west"))) %>% select(-c(docs,beds)) %>% view()
 
@@
   select(-c("skim_type","complete_rate")) %>% 
+    mutate(skim_variable = 
+             recode(skim_variable, pcincome = "pcincome (in dollars)", totalinc = "totalinc (in million of dollars)"
+  )) %>% 
   knitr::kable(
     col.names = c("variable", "n_missing", "mean","sd","min","Q25","median","Q75","max","histogram"),
-    caption = "Global Summary", digits = 2)
+    caption = "Global Summary", digits = 4)
 ```
@@
 ```{r}
-par(mfrow=c(3,5))
+par(mfrow = c(3,5))
 
+boxplot(cdi$density_pop,  main = "density_pop")
 boxplot(cdi$area, main = "area")
+
 boxplot(cdi$pop, main = "pop")
@@
 
-boxplot(cdi$docs, main = "docs")
-boxplot(cdi$beds, main = "beds")
-boxplot(cdi$crimes, main = "crimes")
+boxplot(cdi$pdocs, main = "pdocs")
+boxplot(cdi$pbeds, main = "pbeds")
+
 boxplot(cdi$crm_1000,  main = "crm_1000")
@@
 
-* Here is the correlation with pairs function. More specific than the heat map above. We can observe the correlation bewteen all the terms here
+* Here is the correlation with pairs function. More specific than the heat map above. We can observe the correlation between all the terms here
 ```{r}
@@
 #### Marginal distribution ?
+
+```{r}
+cdi %>% ggplot(aes(x = density_pop, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = area, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = pop, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = pop18, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+# positive correlation
+cdi %>% ggplot(aes(x = pop65, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = pdocs, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = pbeds, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = hsgrad, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red') #negative correlation
+
+cdi %>% ggplot(aes(x = bagrad, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = poverty, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red') # positive correlation
+
+cdi %>% ggplot(aes(x = unemp, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = pcincome, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = totalinc, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+
+cdi %>% ggplot(aes(x = region, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+```
 

### 2021-12-13 draft_yma.Rmd
change to github document for easier access
marginal plot

--- 2021-12-12 draft_yma.Rmd
@@
 date: "12/12/2021"
-output: html_document
+output: github_document
 ---
@@
   mutate(crm_1000 = 1000*(crimes/pop),
+         pdocs_1000 = 1000*(docs/pop),
+         pbeds_1000 = 1000*(beds/pop),
+         density_pop = pop/area,
          # convert region to factors and recoded them accordingly 
          region = factor(region, levels = 1:4,
-                    labels = c("northeast", "north central", "south", "west"))) %>% view()
+                    labels = c("northeast", "north central", "south", "west"))) %>% select(-c(docs,beds)) %>% view()
 
@@
 
-boxplot(cdi$docs, main = "docs")
-boxplot(cdi$beds, main = "beds")
-boxplot(cdi$crimes, main = "crimes")
+boxplot(cdi$pdocs_1000, main = "pdocs_1000")
+boxplot(cdi$pbeds_1000, main = "pbeds_1000")
+
 boxplot(cdi$crm_1000,  main = "crm_1000")
@@
 #### Marginal distribution ?
+
+```{r}
+library(ggplot2)
+library(ggExtra)
+
+```
+
+```{r}
+marg_den = cdi %>% ggplot(aes(x = density_pop, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_den, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_area = cdi %>% ggplot(aes(x = area, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_area, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_pop = cdi %>% ggplot(aes(x = pop, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_pop, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_pop18 = cdi %>% ggplot(aes(x = pop18, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+# positive correlation
+ggMarginal(marg_pop18, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_pop65 = cdi %>% ggplot(aes(x = pop65, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_pop65, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_pdocs_1000 = cdi %>% ggplot(aes(x = pdocs_1000, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_pdocs_1000, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_pbeds_1000 = cdi %>% ggplot(aes(x = pbeds_1000, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_pbeds_1000, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_hsgrad = cdi %>% ggplot(aes(x = hsgrad, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red') #negative correlation
+ggMarginal(marg_hsgrad, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_bagrad = cdi %>% ggplot(aes(x = bagrad, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_bagrad, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_poverty = cdi %>% ggplot(aes(x = poverty, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red') # positive correlation
+ggMarginal(marg_poverty, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_unemp = cdi %>% ggplot(aes(x = unemp, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_unemp, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_pcincome = cdi %>% ggplot(aes(x = pcincome, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_pcincome, type = "histogram", fill="transparent")
+```
+
+```{r}
+marg_totalinc = cdi %>% ggplot(aes(x = totalinc, y = crm_1000)) + geom_point(alpha = 0.3) + geom_smooth(method = 'lm', se = TRUE, color = 'red')
+ggMarginal(marg_totalinc, type = "histogram", fill="transparent")
+```
 

### 2021-12-13 draft_yma.Rmd
first few model

@@
 library(corrplot)
-corrplot(cor(cdi_descriptive), type = "upper", diag = FALSE)
+cor(cdi_descriptive) %>% knitr::kable()
+
+library(ggcorrplot)
+library(ggstatsplot)
+ggstatsplot::ggcorrmat(
+  data = cdi,
+  type = "parametric", # parametric for Pearson, nonparametric for Spearman's correlation
+  colors = c("darkred", "white", "steelblue") # change default colors
+)
 ```
@@
 ```
+
+**we probably need to plot region as well? not sure**
+
+## Model!
+
+**Q: do we need more descriptive analysis, visualization? rate of crime for each state?**
+**there's a few duplicated cty, but different area, maybe diff place then? not sure**
+they can't be the same... otherwise we have to recalculate some stuff...
+duplicated(cdi[,2:3])
+cdi %>% filter(cty == "Baltimor")
+cdi %>% filter(cty == "St._Loui")
+
+#### Full model predictors ok?
+
+included pop in addition to pop18, pop65. Do we need pop tho? Would pop18 and pop65 be enough already?
+
+did not include cty, state, area, crimes, totalinc, totalinc (included pcincome), do we need to include any of those? for example, both totalinc and pcincome, area and pop_den
+
+this model used `northeast` as the reference level for region
+
+```{r}
+cdi_model = cdi %>% select(-c(id,cty,state,area,crimes,totalinc))
+
+# use 
+full_fit = lm(crm_1000 ~ ., data = cdi_model)
+summary(full_fit)
+```
+
+#### Backward
+
+```{r}
+fit_back = step(full_fit, direction='backward')
+fit_back
+```
+
+crm_1000 ~ pop + pop18 + hsgrad + bagrad + poverty + 
+    pcincome + region + pbeds_1000 + density_pop, data = cdi_model
+
+#### forward
+
+```{r}
+fit_forward = step(full_fit, direction='forward')
+fit_forward
+```
+
+crm_1000 ~ pop + pop18 + pop65 + hsgrad + bagrad + 
+    poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
+    density_pop, data = cdi_model
+
+#### both
+
+step-wise?
+
+```{r}
+fit_both = step(full_fit, direction='both')
+fit_both
+```
+
+crm_1000 ~ pop + pop18 + hsgrad + bagrad + poverty + 
+    pcincome + region + pbeds_1000 + density_pop, data = cdi_model
+This gives back the same model as back
+
+Some group forward again based on their first time step wise...not sure if we need to do that tho
+
+### Test based procedures
+
+#### Cp
+
+I didn't include the full model here
+
+**backward & both**
+
+```{r}
+# printing the 2 best models of each size. For example, the first two lines: print the best 2 models that have 2 variables (including intercept)
+library(leaps)
+Cp_b = leaps(x = model.matrix(fit_back)[,-1],
+      y = cdi_model$crm_1000,
+      nbest = 2,
+      method = "Cp")
+Cp_b
+
+Cp_b %>% faraway::Cpplot()
+```
+
+size is the number of parameters in the model, Cp is Cp
+
+The plot shows the model that satisfy the criteria, number indicates the number of parameters it included. in this case, the 1st to the last.
+
+**forward**
+```{r}
+leaps(x = model.matrix(fit_forward)[,-1],
+      y = cdi_model$crm_1000,
+      nbest = 2,
+      method = "Cp") %>% faraway::Cpplot()
+```
+
+So, multiple model satisfies the Cp criteria here.
+
+#### Adj R2
+
+**backward**
+```{r}
+adjr_b = leaps(x = model.matrix(fit_back)[,-1],
+      y = cdi_model$crm_1000,
+      nbest = 2,
+      method = "adjr2")
+
+adjr_b
+
+adjr_b%>% 
+  faraway::maxadjr(10)
+```
+
+**forward**
+```{r}
+leaps(x = model.matrix(fit_forward)[,-1],
+      y = cdi_model$crm_1000,
+      nbest = 2,
+      method = "adjr2") %>% faraway::maxadjr(10)
+```

### 2021-12-14 draft_yma.Rmd
transformation

@@
     caption = "Global Summary", digits = 4)
+
+ggplot(gather(cdi_descriptive), aes(value)) + 
+    geom_histogram(bins = 13) + 
+    facet_wrap(~key, scales = 'free_x')
 ```
@@
 ggstatsplot::ggcorrmat(
-  data = cdi,
+  data = cdi_descriptive,
   type = "parametric", # parametric for Pearson, nonparametric for Spearman's correlation
@@
 
-included pop in addition to pop18, pop65. Do we need pop tho? Would pop18 and pop65 be enough already?
-
-did not include cty, state, area, crimes, totalinc, totalinc (included pcincome), do we need to include any of those? for example, both totalinc and pcincome, area and pop_den
-
 this model used `northeast` as the reference level for region
@@
 summary(full_fit)
+
+ols_plot_resid_qq(full_fit)
+ols_plot_resid_fit(full_fit)
+```
+
+## Transformation
+```{r}
+lambda = MASS::boxcox(full_fit)
+optlam = lambda$x[which.max(lambda$y)]
+optlam
+```
+
+The lambda from the transformation is 0.5454, so we will try to fit a square root transformation to y
+
+```{r}
+trans_fit = lm(crm_1000^0.5 ~ .,data = cdi_model)
+summary(trans_fit)
+
+ols_plot_resid_fit(trans_fit)
+ols_plot_resid_qq(trans_fit)
+
+lambda_trans = MASS::boxcox(trans_fit)
+optlam_trans = lambda_trans$x[which.max(lambda_trans$y)]
+lambda_trans
 ```
@@
 ```
+
+boxcox
+
+```{r}
+lambda = boxcox(cdi_model$crm_1000~1)
+optlam = lambda$x[which.max(lambda$y)]
+```

### 2021-12-14 draft_yma.Rmd
boxcox transformation

@@
 
-ols_plot_resid_qq(full_fit)
-ols_plot_resid_fit(full_fit)
+olsrr::ols_plot_resid_qq(full_fit)
+olsrr::ols_plot_resid_fit(full_fit)
 ```
@@
 
-ols_plot_resid_fit(trans_fit)
-ols_plot_resid_qq(trans_fit)
+olsrr::ols_plot_resid_fit(trans_fit)
+olsrr::ols_plot_resid_qq(trans_fit)
 
@@
 ```
-
-boxcox
-
-```{r}
-lambda = boxcox(cdi_model$crm_1000~1)
-optlam = lambda$x[which.max(lambda$y)]
-```

### 2021-12-14 draft_yma.Rmd
backward forward

@@
 ```{r}
-fit_back = step(full_fit, direction='backward')
+fit_back = step(trans_fit, direction='backward')
 fit_back
 ```
-
-crm_1000 ~ pop + pop18 + hsgrad + bagrad + poverty + 
-    pcincome + region + pbeds_1000 + density_pop, data = cdi_model
 
@@
 ```{r}
-fit_forward = step(full_fit, direction='forward')
+fit_forward = step(trans_fit, direction='forward')
 fit_forward
 ```
-
-crm_1000 ~ pop + pop18 + pop65 + hsgrad + bagrad + 
-    poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
-    density_pop, data = cdi_model
 
@@
 ```{r}
-fit_both = step(full_fit, direction='both')
+fit_both = step(trans_fit, direction='both')
 fit_both

### 2021-12-14 draft_yma.Rmd
forwar, backward with trans

@@
     caption = "Global Summary", digits = 4)
-
-ggplot(gather(cdi_descriptive), aes(value)) + 
-    geom_histogram(bins = 13) + 
-    facet_wrap(~key, scales = 'free_x')
 ```
@@
 ```{r}
-par(mfrow = c(3,5))
+par(mfrow = c(2,7))
 
@@
   geom_histogram()
-
-# log transfor the outcome
-cdi %>% 
-  ggplot(aes(x = log(crm_1000))) +
-  geom_histogram()
-
 ```
@@
 fit_back
+
+olsrr::ols_plot_resid_fit(fit_back)
+olsrr::ols_plot_resid_qq(fit_back)
+MASS::boxcox(fit_back)
 ```
+
+lm(formula = crm_1000^0.5 ~ pop + pop18 + hsgrad + bagrad + poverty + 
+    pcincome + region + pbeds_1000 + density_pop, data = cdi_model)
 
@@
 fit_forward
+
+olsrr::ols_plot_resid_fit(fit_forward)
+olsrr::ols_plot_resid_qq(fit_forward)
+MASS::boxcox(fit_forward)
 ```
+
+lm(formula = crm_1000^0.5 ~ pop + pop18 + pop65 + hsgrad + bagrad + 
+    poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
+    density_pop, data = cdi_model)
 
@@
 ```
+lm(formula = crm_1000^0.5 ~ pop + pop18 + hsgrad + bagrad + poverty + 
+    pcincome + region + pbeds_1000 + density_pop, data = cdi_model)
 
-crm_1000 ~ pop + pop18 + hsgrad + bagrad + poverty + 
-    pcincome + region + pbeds_1000 + density_pop, data = cdi_model
 This gives back the same model as back
-
-Some group forward again based on their first time step wise...not sure if we need to do that tho
 
@@
 ```
+
+## backward by hand
+
+```{r}
+# No docs
+step1 = update(full_fit, . ~ . -pdocs_1000)
+summary(step1)
+
+# No pop65
+step2 = update(step1, . ~ . -pop65)
+summary(step2)
+
+# No unemp
+step3 = update(step2, . ~ . -unemp)
+summary(step3)
+
+mult.fit.final = lm(formula = crm_1000 ~ pop + pop18 + hsgrad + bagrad + poverty + 
+    pcincome + region + pbeds_1000 + density_pop, data = cdi_model)
+summary(mult.fit.final)
+```

### 2021-12-14 draft_yma.Rmd
change the backward bu hand

@@
 # No docs
-step1 = update(full_fit, . ~ . -pdocs_1000)
+step1 = update(trans_fit, . ~ . -pdocs_1000)
 summary(step1)
@@
 
-mult.fit.final = lm(formula = crm_1000 ~ pop + pop18 + hsgrad + bagrad + poverty + 
+step4 = update(step3, . ~ . -hsgrad)
+summary(step4)
+
+mult.fit.final = lm(crm_1000^0.5 ~ pop + pop18 + bagrad + poverty + 
     pcincome + region + pbeds_1000 + density_pop, data = cdi_model)

### 2021-12-14 draft_yma.Rmd
calculated Cp, BIC

--- 2021-12-14 draft_yma.Rmd
@@
 
-```{r}
+```{r boxplot}
 par(mfrow = c(2,7))
@@
 **Q: do we need more descriptive analysis, visualization? rate of crime for each state?**
-**there's a few duplicated cty, but different area, maybe diff place then? not sure**
-they can't be the same... otherwise we have to recalculate some stuff...
-duplicated(cdi[,2:3])
-cdi %>% filter(cty == "Baltimor")
-cdi %>% filter(cty == "St._Loui")
 
-#### Full model predictors ok?
+#### Full model predictors
 
@@
 
-The lambda from the transformation is 0.5454, so we will try to fit a square root transformation to y
+The lambda from the transformation is 0.5454, so we will try to fit a square root transformation to Y
 
@@
 lambda_trans
+
+cdi_trans = cdi_model %>% mutate(crm_1000_sqr = crm_1000^0.5) %>% select(-c(crm_1000))
 ```
@@
 
-### Test based procedures
+### Criterion
 
@@
 
-I didn't include the full model here
+```{r}
+olsrr::ols_mallows_cp(fit_back,trans_fit)
+olsrr::ols_mallows_cp(fit_forward,trans_fit)
+olsrr::ols_mallows_cp(fit_both,trans_fit)
+```
 
-**backward & both**
+#### Adj R2
 
 ```{r}
+summary(fit_back) # 0.5469
+summary(fit_forward) #0.5457
+summary(fit_both) #5469
+```
+
+### BIC
+
+```{r}
+library(flexmix)
+BIC(fit_back)
+BIC(fit_forward)
+BIC(fit_both)
+```
+
+## Cp
+
+didn't use any of this!
+```{r, include=FALSE}
 # printing the 2 best models of each size. For example, the first two lines: print the best 2 models that have 2 variables (including intercept)
 library(leaps)
-Cp_b = leaps(x = model.matrix(fit_back)[,-1],
-      y = cdi_model$crm_1000,
+Cp_b = leaps(x = model.matrix(trans_fit)[,-1],
+      y = cdi_trans$crm_1000_sqr,
       nbest = 2,
       method = "Cp")
-Cp_b
 
-Cp_b %>% faraway::Cpplot()
+mod_b = Cp_b$Cp[which(Cp_b$Cp <= Cp_b$size)]
 ```
@@
 **forward**
-```{r}
-leaps(x = model.matrix(fit_forward)[,-1],
-      y = cdi_model$crm_1000,
+```{r, include=FALSE}
+Cp_f = leaps(x = model.matrix(fit_forward)[,-1],
+      y = cdi_trans$crm_1000_sqr,
       nbest = 2,
-      method = "Cp") %>% faraway::Cpplot()
+      method = "Cp")
+
+mod_f = Cp_f$Cp[which(Cp_f$Cp <= Cp_f$size)]
 ```
@@
 
+Did not use!!
 **backward**
-```{r}
+```{r, include=FALSE}
 adjr_b = leaps(x = model.matrix(fit_back)[,-1],
-      y = cdi_model$crm_1000,
+      y = cdi_trans$crm_1000_sqr,
       nbest = 2,
@@
 ```
-
 **forward**
-```{r}
+```{r, include=FALSE}
 leaps(x = model.matrix(fit_forward)[,-1],
-      y = cdi_model$crm_1000,
+      y = cdi_trans$crm_1000_sqr,
       nbest = 2,

### 2021-12-16 draft_yma.Rmd
all code should be here, discuss tmr

@@
 library(tidyverse)
+library(leaps)
 library(readr)
@@
          region = factor(region, levels = 1:4,
-                    labels = c("northeast", "north central", "south", "west"))) %>% select(-c(docs,beds)) %>% view()
+                    labels = c("northeast", "northcentral", "south", "west"))) %>% select(-c(docs,beds))
 
@@
 
-### States with unusual rates
-
-group by state
-```{r}
-cdi_by_state = cdi %>% group_by(state) %>% 
-  summarise(state_pop = sum(pop),
-            state_crimes = sum(crimes)) %>% 
-  mutate(state_CRM_1000 = 1000*state_crimes/state_pop)
-```
-
-Not sure what counts as out lier, this is what the other group did.
+### Counties with unusual rates
 
@@
 ```
-
-This shows no outlier on the state level tho...
-
-```{r}
-upper = quantile(cdi_by_state$state_CRM_1000, 0.75)
-lower = quantile(cdi_by_state$state_CRM_1000, 0.25)
-IQR = upper - lower
-cdi_by_state %>% 
-  filter(state_CRM_1000 > upper + 1.5*IQR,
-         state_CRM_1000 > lower - 1.5*IQR) %>% 
-  dplyr::select(state, state_CRM_1000) %>%
-  knitr::kable(digits = 2)
-boxplot(cdi_by_state$state_CRM_1000, main = 'State Crime Rate per 1000 people')
-```
-
-**we probably need to plot region as well? not sure**
 
@@
 
+### New tibble with transformed crm_1000 -> crm_1000_sqr
 ```{r}
-trans_fit = lm(crm_1000^0.5 ~ .,data = cdi_model)
+cdi_trans = cdi_model %>% mutate(crm_1000_sqr = crm_1000^0.5) %>% select(-c(crm_1000))
+```
+
+```{r}
+trans_fit = lm(crm_1000_sqr ~ .,data = cdi_trans)
 summary(trans_fit)
@@
 optlam_trans = lambda_trans$x[which.max(lambda_trans$y)]
-lambda_trans
+optlam_trans
+```
 
-cdi_trans = cdi_model %>% mutate(crm_1000_sqr = crm_1000^0.5) %>% select(-c(crm_1000))
+### Counties with unusual rates
+
+```{r}
+upper = quantile(cdi_trans$crm_1000_sqr, 0.75)
+lower = quantile(cdi_trans$crm_1000_sqr, 0.25)
+IQR = upper - lower
+
+cdi_trans %>% 
+  filter(crm_1000_sqr > upper + 1.5*IQR,
+         crm_1000_sqr > lower - 1.5*IQR) %>% 
+  dplyr::select(crm_1000_sqr) %>% 
+  mutate(cty = c("Kings", "St._Loui")) %>% 
+  knitr::kable(digits = 2)
+
+# cdi %>% filter(pop == 2300664 | pop == 396685) %>% select(cty,pop)
 ```
@@
 
-lm(formula = crm_1000^0.5 ~ pop + pop18 + hsgrad + bagrad + poverty + 
-    pcincome + region + pbeds_1000 + density_pop, data = cdi_model)
+Backward: crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop
 
@@
 ```{r}
-fit_forward = step(trans_fit, direction='forward')
+fit_forward = step(trans_fit, direction="forward")
 fit_forward
@@
 
-lm(formula = crm_1000^0.5 ~ pop + pop18 + pop65 + hsgrad + bagrad + 
+lm(formula = crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + 
     poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
-    density_pop, data = cdi_model)
+    density_pop, data = cdi_trans)
 
@@
 
+## Criterion Based: Adjr2 and Cp
+
+```{r}
+# printing the 2 best models of each size. For example, the first two lines: print the best 2 models that have 2 variables (including intercept)
+
+models <- regsubsets(crm_1000_sqr ~., data = cdi_trans,nvmax = 12)
+res_sum = summary(models)
+
+par(mfrow=c(1,2))
+plot(1:12, res_sum$cp, xlab="No of parameters", ylab="Cp Statistic")
+plot(1:12, res_sum$adjr2, xlab="No of parameters", ylab="Adj R2")
+
+mod_s = data.frame(
+  Adj.R2 = which.max(res_sum$adjr2),
+  CP = which.min(res_sum$cp)
+)
+
+get_model_formula <- function(id, object, outcome){
+  # get models data
+  models <- summary(object)$which[id,-1]
+  # Get outcome variable
+  #form <- as.formula(object$call[[2]])
+  #outcome <- all.vars(form)[1]
+  # Get model predictors
+  predictors <- names(which(models == TRUE))
+  predictors <- paste(predictors, collapse = "+")
+  # Build model formula
+  as.formula(paste0(outcome, "~", predictors))
+}
+```
+
+Model chosen by Adj $R^2$:
+
+```{r}
+get_model_formula(12, models, "crm_1000_sqr")
+```
+
+```{r}
+fit_adjr = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = cdi_trans)
+
+olsrr::ols_plot_resid_fit(fit_adjr)
+olsrr::ols_plot_resid_qq(fit_adjr)
+MASS::boxcox(fit_adjr)
+```
+
+crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop
+
+Model chosen by Cp: Same with backward
+
+```{r}
+get_model_formula(11, models, "crm_1000_sqr")
+```
+
+crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region+ pbeds_1000 + density_pop
+
+```{r}
+fit_Cp = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region+ pbeds_1000 + density_pop, data=cdi_trans)
+
+olsrr::ols_plot_resid_fit(fit_Cp)
+olsrr::ols_plot_resid_qq(fit_Cp)
+MASS::boxcox(fit_Cp)
+```
+
 ### Criterion
@@
 olsrr::ols_mallows_cp(fit_both,trans_fit)
+olsrr::ols_mallows_cp(fit_adjr,trans_fit)
+olsrr::ols_mallows_cp(fit_Cp,trans_fit)
 ```
@@
 summary(fit_forward) #0.5457
-summary(fit_both) #5469
+summary(fit_both) #0.5469
+summary(fit_adjr) #0.5474
+summary(fit_Cp) #0.5469
 ```
 
-### BIC
+#### BIC
 
@@
 library(flexmix)
-BIC(fit_back)
-BIC(fit_forward)
-BIC(fit_both)
+BIC(fit_back) #1453.527
+BIC(fit_forward) #1469.908
+BIC(fit_both) #1453.527
+BIC(fit_adjr) # 1458.104
+BIC(fit_Cp) #1453.527
 ```
 
-## Cp
+#### MSE
 
-didn't use any of this!
-```{r, include=FALSE}
-# printing the 2 best models of each size. For example, the first two lines: print the best 2 models that have 2 variables (including intercept)
-library(leaps)
-Cp_b = leaps(x = model.matrix(trans_fit)[,-1],
-      y = cdi_trans$crm_1000_sqr,
-      nbest = 2,
-      method = "Cp")
+```{r}
+back_summ =summary(fit_back) #1.331
+mean(back_summ$residuals^2)
 
-mod_b = Cp_b$Cp[which(Cp_b$Cp <= Cp_b$size)]
+forward_summ =summary(fit_forward) #1.325
+mean(forward_summ$residuals^2)
+
+both_summ =summary(fit_both) #1.331
+mean(both_summ$residuals^2)
+
+adjr_summ =summary(fit_adjr) #1.326
+mean(adjr_summ$residuals^2)
+
+Cp_summ =summary(fit_Cp) #1.331
+mean(Cp_summ$residuals^2)
 ```
 
-size is the number of parameters in the model, Cp is Cp
+## Check Collinearity
 
-The plot shows the model that satisfy the criteria, number indicates the number of parameters it included. in this case, the 1st to the last.
-
-**forward**
-```{r, include=FALSE}
-Cp_f = leaps(x = model.matrix(fit_forward)[,-1],
-      y = cdi_trans$crm_1000_sqr,
-      nbest = 2,
-      method = "Cp")
-
-mod_f = Cp_f$Cp[which(Cp_f$Cp <= Cp_f$size)]
+```{r}
+performance::check_collinearity(fit_back)
+performance::check_collinearity(fit_forward)
+performance::check_collinearity(fit_adjr)
+performance::check_collinearity(fit_Cp)
 ```
 
-So, multiple model satisfies the Cp criteria here.
+delete vif larger than 10. Non of them are greater than 10. So all of the models are good, don't have to worry about collinearity.
 
-#### Adj R2
+```{r}
+par(mfrow=c(2,2))
+plot(fit_back)
+```
 
-Did not use!!
-**backward**
-```{r, include=FALSE}
-adjr_b = leaps(x = model.matrix(fit_back)[,-1],
-      y = cdi_trans$crm_1000_sqr,
-      nbest = 2,
-      method = "adjr2")
+## Checking to Outliers and Influential Points
 
-adjr_b
+### Back
 
-adjr_b%>% 
-  faraway::maxadjr(10)
+#### Old Method
+```{r}
+# residuals vs leverage plot
+plot(fit_back, which = 4)
+#remove outliers
+cdi_trans_without = cdi_trans[-c(1, 6, 53),]
+fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + 
+    pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
+summary(fit_without)
 ```
-**forward**
-```{r, include=FALSE}
-leaps(x = model.matrix(fit_forward)[,-1],
-      y = cdi_trans$crm_1000_sqr,
-      nbest = 2,
-      method = "adjr2") %>% faraway::maxadjr(10)
+
+```{r}
+par(mfrow=c(2,2))
+plot(fit_without)
+```
+
+### Threshold: $D_{i}>\frac{4}{n} = \frac{4}{440} = 0.009$
+```{r}
+cook = olsrr::ols_plot_cooksd_bar(fit_back, print_plot = TRUE)
+normal = cook$data$color == "normal"
+outlier = cook$data$color == "outlier"
+```
+
+```{r}
+outlier_df = cdi[outlier,]
+```
+
+This gives us 21 outliers. We should consider if we want to take them all out
+
+### Build model without outlier
+
+Here's what happens when we take them all out.
+
+```{r}
+cdi_trans_without = cdi_trans[normal,]
+```
+
+Built the without model with **fit_back**
+
+```{r}
+fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
+
+summary(fit_without)
+
+# Residual performance
+par(mfrow=c(2,2))
+olsrr::ols_plot_resid_fit(fit_without)
+olsrr::ols_plot_resid_qq(fit_without)
+MASS::boxcox(fit_without)
+
+#Cp -88.45??
+olsrr::ols_mallows_cp(fit_without,trans_fit)
+#BIC: 1274.534
+BIC(fit_without) 
+#MSE
+without_summ =summary(fit_without) #1.017
+mean(without_summ$residuals^2)
+#VIF 
+performance::check_collinearity(fit_without)
+```
+
+Built the without model with **fit_adjr**
+
+```{r}
+fit_without2 = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = cdi_trans_without)
+
+summary(fit_without2)
+
+# Residual performance
+par(mfrow=c(2,2))
+olsrr::ols_plot_resid_fit(fit_without2)
+olsrr::ols_plot_resid_qq(fit_without2)
+MASS::boxcox(fit_without2)
+
+#Cp -87.5??
+olsrr::ols_mallows_cp(fit_without2,trans_fit)
+#BIC: 1279.154
+BIC(fit_without2)
+#MSE
+without2_summ =summary(fit_without2) #1.013
+mean(without2_summ$residuals^2)
+#VIF 
+performance::check_collinearity(fit_without2) 
+```
+
+# Model Validation
+
+## Compute RMSE and adjusted R^2 by cross-validation
+
+Using back and adjr model and data without outlier
+
+```{r warning=FALSE}
+library(modelr)
+set.seed(1)
+cv_df = 
+  crossv_kfold(cdi_trans_without, k = 5) %>% 
+  mutate(
+    train = map(train, as_tibble),
+    test = map(test, as_tibble))
+
+cv_df = 
+  cv_df %>%
+  mutate(
+    fit_back = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = .x)),
+    trans_fit = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + pop65+ hsgrad + bagrad + poverty + unemp + pcincome + region+ +pdocs_1000+pbeds_1000 + density_pop, data = .x)),
+    fit_adjr = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x)),
+    fit_forward = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x))) %>% 
+  mutate(
+    rmse_model2 = map2_dbl(fit_back, test, ~rmse(model = .x, data = .y)),
+    rmse_model3 = map2_dbl(trans_fit, test, ~rmse(model = .x, data = .y)),
+    rmse_model4 = map2_dbl(fit_adjr, test, ~rmse(model = .x, data = .y)),
+    rmse_model5 = map2_dbl(fit_forward, test, ~rmse(model = .x, data = .y))) %>%
+  mutate(
+    mspe_model2 = map2_dbl(fit_back, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
+    mspe_model3 = map2_dbl(trans_fit, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
+    mspe_model4 = map2_dbl(fit_adjr, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
+    mspe_model5 = map2_dbl(fit_forward, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2))
+  ) %>% 
+  mutate(
+    res_model2 = map(fit_back, broom::glance %>% as.data.frame),
+    res_model3 = map(trans_fit, broom::glance %>% as.data.frame),
+    res_model4 = map(fit_adjr, broom::glance %>% as.data.frame),
+    res_model5 = map(fit_forward, broom::glance %>% as.data.frame)) %>%
+  unnest(res_model2, res_model3, res_model4, res_model5) %>%
+  dplyr::select(rmse_model2,rmse_model3,rmse_model4,rmse_model5,
+                mspe_model2,mspe_model3,mspe_model4,mspe_model5,
+                value.adj.r.squared,value.adj.r.squared1,value.adj.r.squared2,value.adj.r.squared3) %>%
+  rename(adjR_model2 = value.adj.r.squared,
+         adjR_model3 = value.adj.r.squared1,
+         adjR_model4 = value.adj.r.squared2,
+         adjR_model5 = value.adj.r.squared3)
+
+cv_df %>%
+  summarise_each(funs(mean( .,na.rm = TRUE))) %>%
+  t()
+```
+
+## Plot the violin plot
+```{r warning=FALSE}
+unnest_cd_df = cv_df %>%
+  pivot_longer(rmse_model2:adjR_model5, 
+               names_pattern = "(.*)(......)$", 
+               names_to = c("limit", "model")) %>% 
+  mutate(limit=ifelse(limit=="", "value", limit)) %>%
+  pivot_wider(id_cols = model, 
+              names_from = limit, 
+              values_from = value, 
+              names_repair = "check_unique") %>% 
+  unnest(c(rmse_, adjR_)) %>%
+  rename(rmse = rmse_,
+         adjR = adjR_)
+unnest_cd_df %>% ggplot(aes(x = model, y = rmse)) + geom_violin() 
+unnest_cd_df %>% ggplot(aes(x = model, y = adjR)) + geom_violin() 
 ```

### 2021-12-16 ref.Rmd
all code should be here, discuss tmr

---
title: "Final Project"
author: "Jiacheng Wu"
date: "12/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(corrplot)
library(leaps)
library(MASS)
library(performance)
```

```{r}
crime_df = read.csv("cdi.csv")
```

```{r}
crime_df_1 = 
  dplyr::select(crime_df, -id, -cty, -state)

crime_df_1 = 
  crime_df_1 %>% 
  mutate(CRM.1000 = crimes*1000/pop) %>% 
  dplyr::select(-crimes)
```

```{r}
pairs(crime_df_1)
```

```{r}
corrplot(cor(crime_df_1), type = "upper", diag = FALSE)
```

```{r}
boxplot(crime_df_1$pop, main = "Total population")
boxplot(crime_df_1$pop18, main = "Percent of population aged 18-34")
boxplot(crime_df_1$pop65, main = "Percent of population aged 65+")
boxplot(crime_df_1$area, main = "Land area")
boxplot(crime_df_1$docs, main = "Number of active physicians")
boxplot(crime_df_1$beds, main = "Number of hospital beds")
boxplot(crime_df_1$CRM.1000, main = "Total serious crimes")
boxplot(crime_df_1$hsgrad, main = "Percent high school graduates")
boxplot(crime_df_1$bagrad, main = "Percent bachelor's degrees")
boxplot(crime_df_1$poverty, main = "Percent below poverty level")
boxplot(crime_df_1$unemp, main = "Percent unemployment")
boxplot(crime_df_1$pcincome, main = "Per capita income")
boxplot(crime_df_1$totalinc, main = "Total personal income")
boxplot(crime_df_1$region, main = "Geographical region")

mult.fit1 = lm(CRM.1000 ~ ., data = crime_df_1)
summary(mult.fit1)
```

## Model diagnostics

## Residual vs Fitted & QQ Plots
```{r}

# residual vs fitted plot
plot(mult.fit1, which = 1)

# QQ plot
plot(mult.fit1, which = 2)

# read new data with Ln transformation
crime_df = read.csv("cdi.csv")
crime_df_2 = 
  dplyr::select(crime_df, -id, -cty, -state)

# fit model - log transform the outcome
mult.fit2 = lm(log(CRM.1000) ~ ., data = crime_df_1)

# residual vs fitted plot
plot(mult.fit2, which = 1)

# QQ plot
plot(mult.fit2, which = 2)

```

## Backwards Elimination

Take out non-significant variables one at a time starting with the highest p-value.

```{r}

step1 = update(mult.fit1, . ~ . -hsgrad)
summary(step1)

step2 = update(step1, . ~ . -unemp)
summary(step2)

step3 = update(step2, . ~ . -bagrad)
summary(step3)

step4 = update(step3, . ~ . -docs)
summary(step4)

step5 = update(step4, . ~ . -pop65)
summary(step5)

step(mult.fit1, direction='backward')
step
```

## Forward Selection

Reversed backward elimination starting with the lowest p-value.
### Step 1:  Fit simple linear regressions for all variables,look for the variable with lowest p-value
```{r}

fit1 = lm(CRM.1000 ~ area, data = crime_df_1)
summary(fit1)
fit2 = lm(CRM.1000 ~ pop, data = crime_df_1)
summary(fit2)
fit3 = lm(CRM.1000 ~ pop18, data = crime_df_1)
summary(fit3)
fit4 = lm(CRM.1000 ~ pop65, data = crime_df_1)
summary(fit4)
fit5 = lm(CRM.1000 ~ docs, data = crime_df_1)
summary(fit5)
fit6 = lm(CRM.1000 ~ beds, data = crime_df_1)
summary(fit6)
fit7 = lm(CRM.1000 ~ hsgrad, data = crime_df_1)
summary(fit7)
fit8 = lm(CRM.1000 ~ bagrad, data = crime_df_1)
summary(fit8)
fit9 = lm(CRM.1000 ~ poverty, data = crime_df_1)
summary(fit9)
fit10 = lm(CRM.1000 ~ unemp, data = crime_df_1)
summary(fit10)
fit11 = lm(CRM.1000 ~ pcincome, data = crime_df_1)
summary(fit11)
fit12 = lm(CRM.1000 ~ totalinc, data = crime_df_1)
summary(fit12)
fit13 = lm(CRM.1000 ~ region, data = crime_df_1)
summary(fit13)
```
# Enter first the one with the lowest p-value: poverty
```{r}
forward1 = lm(CRM.1000 ~ poverty, data = crime_df_1)
summary(forward1)

### Step 2: Enter the one with the lowest p-value in the rest 
fit1 = update(forward1, . ~ . +area)
summary(fit1)
fit2 = update(forward1, . ~ . +pop)
summary(fit2)
fit3 = update(forward1, . ~ . +pop18)
summary(fit3)
fit4 = update(forward1, . ~ . +pop65)
summary(fit4)
fit5 = update(forward1, . ~ . +docs)
summary(fit5)
fit6 = update(forward1, . ~ . +beds)
summary(fit6)
fit7 = update(forward1, . ~ . +hsgrad)
summary(fit7)
fit8 = update(forward1, . ~ . +bagrad)
summary(fit8)
fit9 = update(forward1, . ~ . +unemp)
summary(fit9)
fit10 = update(forward1, . ~ . +pcincome)
summary(fit10)
fit11 = update(forward1, . ~ . +totalinc)
summary(fit11)
fit12 = update(forward1, . ~ . +region)
summary(fit12)
```
# Enter the one with the lowest p-value: beds
```{r}
forward2 = update(forward1, . ~ . + beds)
summary(forward2)
### Step 3: Enter the one with the lowest p-value in the rest 
fit1 = update(forward2, . ~ . +area)
summary(fit1)
fit2 = update(forward2, . ~ . +pop)
summary(fit2)
fit3 = update(forward2, . ~ . +pop18)
summary(fit3)
fit4 = update(forward2, . ~ . +pop65)
summary(fit4)
fit5 = update(forward2, . ~ . +docs)
summary(fit5)
fit6 = update(forward2, . ~ . +hsgrad)
summary(fit6)
fit7 = update(forward2, . ~ . +bagrad)
summary(fit7)
fit8 = update(forward2, . ~ . +unemp)
summary(fit8)
fit9 = update(forward2, . ~ . +pcincome)
summary(fit9)
fit10 = update(forward2, . ~ . +totalinc)
summary(fit10)
fit11 = update(forward2, . ~ . +region)
summary(fit11)
```

# Enter the one with the lowest p-value: region

```{r}
forward3 = update(forward2, . ~ . + region)
summary(forward3)
### Step 4: Enter the one with the lowest p-value in the rest 
fit1 = update(forward3, . ~ . +area)
summary(fit1)
fit2 = update(forward3, . ~ . +pop)
summary(fit2)
fit3 = update(forward3, . ~ . +pop18)
summary(fit3)
fit4 = update(forward3, . ~ . +pop65)
summary(fit4)
fit5 = update(forward3, . ~ . +docs)
summary(fit5)
fit6 = update(forward3, . ~ . +hsgrad)
summary(fit6)
fit7 = update(forward3, . ~ . +bagrad)
summary(fit7)
fit8 = update(forward3, . ~ . +unemp)
summary(fit8)
fit9 = update(forward3, . ~ . +pcincome)
summary(fit9)
fit10 = update(forward3, . ~ . +totalinc)
summary(fit10)
```
# Enter the one with the lowest p-value: totalinc
```{r}
forward4 = update(forward3, . ~ . + totalinc)
summary(forward4)
### Step 4: Enter the one with the lowest p-value in the rest 
fit1 = update(forward4, . ~ . +area)
summary(fit1)
fit2 = update(forward4, . ~ . +pop)
summary(fit2)
fit3 = update(forward4, . ~ . +pop18)
summary(fit3)
fit4 = update(forward4, . ~ . +pop65)
summary(fit4)
fit5 = update(forward4, . ~ . +docs)
summary(fit5)
fit6 = update(forward4, . ~ . +hsgrad)
summary(fit6)
fit7 = update(forward4, . ~ . +bagrad)
summary(fit7)
fit8 = update(forward4, . ~ . +unemp)
summary(fit8)
fit9 = update(forward4, . ~ . +pcincome)
summary(fit9)
```
# Enter the one with the lowest p-value: pcincome
```{r}
forward5 = update(forward4, . ~ . + pcincome)
summary(forward4)
### Step 5: Enter the one with the lowest p-value in the rest 
fit1 = update(forward5, . ~ . +area)
summary(fit1)
fit2 = update(forward5, . ~ . +pop)
summary(fit2)
fit3 = update(forward5, . ~ . +pop18)
summary(fit3)
fit4 = update(forward5, . ~ . +pop65)
summary(fit4)
fit5 = update(forward5, . ~ . +docs)
summary(fit5)
fit6 = update(forward5, . ~ . +hsgrad)
summary(fit6)
fit7 = update(forward5, . ~ . +bagrad)
summary(fit7)
fit8 = update(forward5, . ~ . +unemp)
summary(fit8)
```
# Enter the one with the lowest p-value: pop
```{r}
forward6 = update(forward5, . ~ . + pop)
summary(forward5)
### Step 6: Enter the one with the lowest p-value in the rest 
fit1 = update(forward6, . ~ . +area)
summary(fit1)
fit2 = update(forward6, . ~ . +pop18)
summary(fit2)
fit3 = update(forward6, . ~ . +pop65)
summary(fit3)
fit4 = update(forward6, . ~ . +docs)
summary(fit4)
fit5 = update(forward6, . ~ . +hsgrad)
summary(fit5)
fit6 = update(forward6, . ~ . +bagrad)
summary(fit6)
fit7 = update(forward6, . ~ . +unemp)
summary(fit7)
```

# Enter the one with the lowest p-value: pop18

```{r}
forward7 = update(forward6, . ~ . + pop18)
summary(forward7)
### Step 6: Enter the one with the lowest p-value in the rest 
fit1 = update(forward7, . ~ . +area)
summary(fit1)
fit2 = update(forward7, . ~ . +pop65)
summary(fit2)
fit3 = update(forward7, . ~ . +docs)
summary(fit3)
fit4 = update(forward7, . ~ . +hsgrad)
summary(fit4)
fit5 = update(forward7, . ~ . +bagrad)
summary(fit5)
fit6 = update(forward7, . ~ . +unemp)
summary(fit6)
```

# Enter the one with the lowest p-value:area 
```{r}
forward8 = update(forward7, . ~ . + area)
summary(forward8)
### Step 8: Enter the one with the lowest p-value in the rest 
fit1 = update(forward8, . ~ . +pop65)
summary(fit1)
fit2 = update(forward8, . ~ . +docs)
summary(fit2)
fit3 = update(forward8, . ~ . +hsgrad)
summary(fit3)
fit4 = update(forward8, . ~ . +bagrad)
summary(fit4)
fit5 = update(forward8, . ~ . +unemp)
summary(fit5)
```

# P-value of all new added variables are larger than 0.05, which means that they 
# are not significant predictor, and we stop here.

```{r}
# The model we obtained is CRM.1000 ~ poverty + beds + region + totalinc + pcincome + pop + pop18 + area
mult.fit1.final = lm(CRM.1000 ~ poverty + beds + region + totalinc + pcincome + pop + pop18 + area, data = crime_df_1)
summary(mult.fit1.final)

# fit using one function
step(mult.fit1, direction = 'forward')
```

## Stepwise

```{r}
step.model = stepAIC(mult.fit1, direction = "both", 
                      trace = FALSE)
summary(step.model)
```

## Criterion based

```{r}
mat = as.matrix(crime_df_1)

leaps(x = mat[,1:13], y = mat[,14], nbest = 1, method = "Cp")

leaps(x = mat[,1:13], y = mat[,14], nbest = 1, method = "adjr2")

# Function regsubsets() performs a subset selection by identifying the "best" model that contains
# a certain number of predictors. By default "best" is chosen using SSE/RSS (smaller is better)
b = regsubsets(CRM.1000 ~ ., data = crime_df_1)
rs = summary(b)

# plot of Cp and Adj-R2 as functions of parameters
par(mfrow=c(1,2))

plot(1:8, rs$cp, xlab="No of parameters", ylab="Cp Statistic")
abline(0,1)

plot(1:8, rs$adjr2, xlab="No of parameters", ylab="Adj R2")
```

## Diagnosing stepwise model

```{r}
plot(step.model, which = 1)
plot(step.model, which = 2)

# fit model - log transform the outcome
step.modellog = lm(log(CRM.1000) ~ ., data = crime_df_1)

# residual vs fitted plot
plot(step.modellog, which = 1)

# QQ plot
plot(step.modellog, which = 2)

```

```{r}
boxcox(step.model, lambda = seq(-3, 3, by = 0.25))
step.modelsqrt = lm(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + poverty + pcincome + totalinc + region, data = crime_df_1)

boxcox(step.modelsqrt)
```

```{r}
plot(step.model, which = 4)

crimeOut = crime_df_1[-c(1,6),]

plot(crime_df_1$pcincome, crime_df_1$CRM.1000)
plot(crimeOut$pcincome, crimeOut$CRM.1000)

with = lm(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + 
    poverty + pcincome + totalinc + region, data = crime_df_1)

without = lm(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + 
    poverty + pcincome + totalinc + region, data = crimeOut)

plot(without)
```

```{r}
crime_df_10 = dplyr::select(crime_df_1, -bagrad, -hsgrad, -unemp)

corrplot(cor(crime_df_10), type = "upper", diag = FALSE)

pairs(crime_df_10)

check_collinearity(without)
```

Final model selected by the step-wise process

```{r}
final.step.model = lm(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + poverty + pcincome + totalinc + region, data = crimeOut)

summary(final.step.model)
```

Validation of stepwise before diagnostics and remedies
```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_caret = train(CRM.1000 ~ area + pop + pop18 + pop65 + docs + beds + poverty + pcincome + totalinc + region, data = crime_df_1,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

model_caret$finalModel
print(model_caret)
```

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_caret = train(sqrt(CRM.1000) ~ area + pop + pop18 + pop65 + docs + beds + poverty + pcincome + totalinc + region, data = crimeOut,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

model_caret$finalModel
print(model_caret)
```

```{r}
confint(final.step.model)
confint(final.step.model,level = 0.90)
```

### 2021-12-17 final_project.Rmd
finished code

--- 2021-12-16 draft_yma.Rmd
@@
 ---
-title: "draft_yma"
+title: "Final project"
 author: "Anna Ma"
-date: "12/12/2021"
+date: "12/17/2021"
 output: github_document
@@
 library(readr)
+library(flexmix)
 ```
 
-## Data exploration
-
- **converted variables to per person and docs, beds, etc.**
+# Data exploration
 
@@
 
-### Descriptive statistics of all varaibles
+## Descriptive statistics of all varaibles
 ```{r statistic summary}
@@
 
-Q: 1. do we need to group them by state or county? group by gounty gives 5000+ rows though...
-  2. do we need box plot still? 
-  3. again, transfer some variables to "per pop" / "per 1000 pop" ?
-
-### Descriptive analysis
+## Descriptive analysis
 
@@
 boxplot(cdi$totalinc, main = "totalinc")
-
 ```
 
-### Correlation
+## Correlation
 
-#### Pairwise relationship
+### Pairwise relationship
 
@@
 library(ggcorrplot)
-library(ggstatsplot)
-ggstatsplot::ggcorrmat(
-  data = cdi_descriptive,
-  type = "parametric", # parametric for Pearson, nonparametric for Spearman's correlation
-  colors = c("darkred", "white", "steelblue") # change default colors
-)
+
+ggcorrplot(cor(cdi_descriptive), type = "upper",
+   lab = TRUE)
 ```
 
-* Here is the correlation with pairs function. More specific than the heat map above. We can observe the correlation between all the terms here
+Scatter plot
+
 ```{r}
@@
 
-#### Marginal distribution ?
+## Marginal distribution
 
@@
 library(ggExtra)
-
 ```
@@
 
-### Distribution of outcome
+## Distribution of outcome
 
@@
 
-### Counties with unusual rates
+## Counties with unusual rates
 
@@
 
-## Model!
+# Model
 
-**Q: do we need more descriptive analysis, visualization? rate of crime for each state?**
-
-#### Full model predictors
+## Full model predictors
 
@@
 
-### New tibble with transformed crm_1000 -> crm_1000_sqr
+New tibble with transformed Y:
 ```{r}
@@
   knitr::kable(digits = 2)
+```
 
-# cdi %>% filter(pop == 2300664 | pop == 396685) %>% select(cty,pop)
-```
+## Model Selection
+
+### Step-wise approach
 
@@
 
-lm(formula = crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + 
+model: crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + 
     poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
@@
 
-step-wise?
+step-wise
 
@@
 
-## Criterion Based: Adjr2 and Cp
+### Criterion Based: Adjr2 and Cp
 
@@
 
-### Criterion
-
-#### Cp
-
-```{r}
-olsrr::ols_mallows_cp(fit_back,trans_fit)
-olsrr::ols_mallows_cp(fit_forward,trans_fit)
-olsrr::ols_mallows_cp(fit_both,trans_fit)
-olsrr::ols_mallows_cp(fit_adjr,trans_fit)
-olsrr::ols_mallows_cp(fit_Cp,trans_fit)
-```
-
-#### Adj R2
-
-```{r}
-summary(fit_back) # 0.5469
-summary(fit_forward) #0.5457
-summary(fit_both) #0.5469
-summary(fit_adjr) #0.5474
-summary(fit_Cp) #0.5469
-```
-
-#### BIC
-
-```{r}
-library(flexmix)
-BIC(fit_back) #1453.527
-BIC(fit_forward) #1469.908
-BIC(fit_both) #1453.527
-BIC(fit_adjr) # 1458.104
-BIC(fit_Cp) #1453.527
-```
-
-#### MSE
-
-```{r}
-back_summ =summary(fit_back) #1.331
-mean(back_summ$residuals^2)
-
-forward_summ =summary(fit_forward) #1.325
-mean(forward_summ$residuals^2)
-
-both_summ =summary(fit_both) #1.331
-mean(both_summ$residuals^2)
-
-adjr_summ =summary(fit_adjr) #1.326
-mean(adjr_summ$residuals^2)
-
-Cp_summ =summary(fit_Cp) #1.331
-mean(Cp_summ$residuals^2)
-```
-
 ## Check Collinearity
@@
 
-delete vif larger than 10. Non of them are greater than 10. So all of the models are good, don't have to worry about collinearity.
-
-```{r}
-par(mfrow=c(2,2))
-plot(fit_back)
-```
+Delete predictors when vif larger than 10. Non of them are greater than 10. So all of the models are good, don't have to worry about collinearity.
 
@@
 
-### Back
-
-#### Old Method
 ```{r}
-# residuals vs leverage plot
-plot(fit_back, which = 4)
-#remove outliers
-cdi_trans_without = cdi_trans[-c(1, 6, 53),]
-fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + 
-    pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
-summary(fit_without)
+olsrr::ols_plot_cooksd_bar(fit_back, print_plot = TRUE)
+#remove observations whose cook's D > 0.5
+cdi_trans_without = cdi_trans[-c(1, 6),]
 ```
 
-```{r}
-par(mfrow=c(2,2))
-plot(fit_without)
-```
-
-### Threshold: $D_{i}>\frac{4}{n} = \frac{4}{440} = 0.009$
-```{r}
-cook = olsrr::ols_plot_cooksd_bar(fit_back, print_plot = TRUE)
-normal = cook$data$color == "normal"
-outlier = cook$data$color == "outlier"
-```
-
-```{r}
-outlier_df = cdi[outlier,]
-```
-
-This gives us 21 outliers. We should consider if we want to take them all out
-
-### Build model without outlier
-
-Here's what happens when we take them all out.
-
-```{r}
-cdi_trans_without = cdi_trans[normal,]
-```
+## Build model without outlier
 
@@
 ```{r}
-fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
+fit_back_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
 
-summary(fit_without)
+summary(fit_back_without)
 
@@
 par(mfrow=c(2,2))
-olsrr::ols_plot_resid_fit(fit_without)
-olsrr::ols_plot_resid_qq(fit_without)
-MASS::boxcox(fit_without)
+olsrr::ols_plot_resid_fit(fit_back_without)
+olsrr::ols_plot_resid_qq(fit_back_without)
+```
 
-#Cp -88.45??
-olsrr::ols_mallows_cp(fit_without,trans_fit)
-#BIC: 1274.534
-BIC(fit_without) 
-#MSE
-without_summ =summary(fit_without) #1.017
-mean(without_summ$residuals^2)
-#VIF 
-performance::check_collinearity(fit_without)
+Build the without model with **trans_fit** (full model)
+
+```{r}
+trans_fit_without = lm(crm_1000_sqr ~., data = cdi_trans_without)
+
+summary(trans_fit_without)
+
+# Residual performance
+par(mfrow=c(2,2))
+olsrr::ols_plot_resid_fit(trans_fit_without)
+olsrr::ols_plot_resid_qq(trans_fit_without)
 ```
@@
 ```{r}
-fit_without2 = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = cdi_trans_without)
+fit_adjr_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = cdi_trans_without)
 
-summary(fit_without2)
+summary(fit_adjr_without)
 
@@
 par(mfrow=c(2,2))
-olsrr::ols_plot_resid_fit(fit_without2)
-olsrr::ols_plot_resid_qq(fit_without2)
-MASS::boxcox(fit_without2)
+olsrr::ols_plot_resid_fit(fit_adjr_without)
+olsrr::ols_plot_resid_qq(fit_adjr_without)
+```
 
-#Cp -87.5??
-olsrr::ols_mallows_cp(fit_without2,trans_fit)
-#BIC: 1279.154
-BIC(fit_without2)
-#MSE
-without2_summ =summary(fit_without2) #1.013
-mean(without2_summ$residuals^2)
-#VIF 
-performance::check_collinearity(fit_without2) 
+## Criteria
+
+#### Cp for models fit without outlier
+
+```{r}
+olsrr::ols_mallows_cp(fit_back_without,trans_fit) #-14.65
+olsrr::ols_mallows_cp(trans_fit_without,trans_fit)# -10.59
+olsrr::ols_mallows_cp(fit_adjr_without,trans_fit) #-14.49
+```
+
+#### AIC
+
+```{r}
+AIC(fit_back_without) # 1371.373
+AIC(trans_fit_without) #1375.258
+AIC(fit_adjr_without) # 1371.363
+```
+
+#### BIC
+
+```{r}
+BIC(fit_back_without) #1424.441
+BIC(trans_fit_without) #1440,573
+BIC(fit_adjr_without) # 1428.514
+```
+
+#### MSE
+
+```{r}
+back_without_summ =summary(fit_back_without) #1.2633
+mean(back_without_summ$residuals^2)
+
+trans_without_summ =summary(trans_fit_without) #1.2572
+mean(trans_without_summ$residuals^2)
+
+adjr_without_summ =summary(fit_adjr_without) #1.2575
+mean(adjr_without_summ$residuals^2)
 ```
@@
 
-## Compute RMSE and adjusted R^2 by cross-validation
+## Compute RMSE, MSPE, adjusted R^2 by cross-validation
 
@@
     trans_fit = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + pop65+ hsgrad + bagrad + poverty + unemp + pcincome + region+ +pdocs_1000+pbeds_1000 + density_pop, data = .x)),
-    fit_adjr = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x)),
-    fit_forward = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x))) %>% 
+    fit_adjr = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x))) %>% 
   mutate(
-    rmse_model2 = map2_dbl(fit_back, test, ~rmse(model = .x, data = .y)),
-    rmse_model3 = map2_dbl(trans_fit, test, ~rmse(model = .x, data = .y)),
-    rmse_model4 = map2_dbl(fit_adjr, test, ~rmse(model = .x, data = .y)),
-    rmse_model5 = map2_dbl(fit_forward, test, ~rmse(model = .x, data = .y))) %>%
+    rmse_back = map2_dbl(fit_back, test, ~rmse(model = .x, data = .y)),
+    rmse_full = map2_dbl(trans_fit, test, ~rmse(model = .x, data = .y)),
+    rmse_adjr = map2_dbl(fit_adjr, test, ~rmse(model = .x, data = .y))) %>%
   mutate(
-    mspe_model2 = map2_dbl(fit_back, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
-    mspe_model3 = map2_dbl(trans_fit, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
-    mspe_model4 = map2_dbl(fit_adjr, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
-    mspe_model5 = map2_dbl(fit_forward, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2))
-  ) %>% 
+    mspe_back = map2_dbl(fit_back, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
+    mspe_full = map2_dbl(trans_fit, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
+    mspe_adjr = map2_dbl(fit_adjr, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2))) %>% 
   mutate(
-    res_model2 = map(fit_back, broom::glance %>% as.data.frame),
-    res_model3 = map(trans_fit, broom::glance %>% as.data.frame),
-    res_model4 = map(fit_adjr, broom::glance %>% as.data.frame),
-    res_model5 = map(fit_forward, broom::glance %>% as.data.frame)) %>%
-  unnest(res_model2, res_model3, res_model4, res_model5) %>%
-  dplyr::select(rmse_model2,rmse_model3,rmse_model4,rmse_model5,
-                mspe_model2,mspe_model3,mspe_model4,mspe_model5,
-                value.adj.r.squared,value.adj.r.squared1,value.adj.r.squared2,value.adj.r.squared3) %>%
-  rename(adjR_model2 = value.adj.r.squared,
-         adjR_model3 = value.adj.r.squared1,
-         adjR_model4 = value.adj.r.squared2,
-         adjR_model5 = value.adj.r.squared3)
+    res_back = map(fit_back, broom::glance %>% as.data.frame),
+    res_full = map(trans_fit, broom::glance %>% as.data.frame),
+    res_adjr = map(fit_adjr, broom::glance %>% as.data.frame)) %>%
+  unnest(res_back, res_full, res_adjr) %>%
+  dplyr::select(rmse_back,rmse_full,rmse_adjr,
+                mspe_back,mspe_full,mspe_adjr,
+                value.adj.r.squared,value.adj.r.squared1,value.adj.r.squared2) %>%
+  rename(adjR_back = value.adj.r.squared,
+         adjR_full = value.adj.r.squared1,
+         adjR_adjr = value.adj.r.squared2)
 
@@
 unnest_cd_df = cv_df %>%
-  pivot_longer(rmse_model2:adjR_model5, 
-               names_pattern = "(.*)(......)$", 
+  pivot_longer(rmse_back:adjR_adjr, 
+               names_pattern = "(.*)(....)$", 
                names_to = c("limit", "model")) %>% 
@@
 ```
-
-## backward by hand
-
-```{r}
-# No docs
-step1 = update(trans_fit, . ~ . -pdocs_1000)
-summary(step1)
-
-# No pop65
-step2 = update(step1, . ~ . -pop65)
-summary(step2)
-
-# No unemp
-step3 = update(step2, . ~ . -unemp)
-summary(step3)
-
-step4 = update(step3, . ~ . -hsgrad)
-summary(step4)
-
-mult.fit.final = lm(crm_1000^0.5 ~ pop + pop18 + bagrad + poverty + 
-    pcincome + region + pbeds_1000 + density_pop, data = cdi_model)
-summary(mult.fit.final)
-```

### 2021-12-17 final_project.tex
finished code

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\title{Final project}
\author{Anna Ma}
\date{12/17/2021}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Final project},
  pdfauthor={Anna Ma},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\begin{document}
\maketitle

\hypertarget{data-exploration}{%
\section{Data exploration}\label{data-exploration}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cdi }\OtherTok{=} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"cdi.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  janitor}\SpecialCharTok{::}\FunctionTok{clean\_names}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{crm\_1000 =} \DecValTok{1000}\SpecialCharTok{*}\NormalTok{(crimes}\SpecialCharTok{/}\NormalTok{pop),}
         \AttributeTok{pdocs\_1000 =} \DecValTok{1000}\SpecialCharTok{*}\NormalTok{(docs}\SpecialCharTok{/}\NormalTok{pop),}
         \AttributeTok{pbeds\_1000 =} \DecValTok{1000}\SpecialCharTok{*}\NormalTok{(beds}\SpecialCharTok{/}\NormalTok{pop),}
         \AttributeTok{density\_pop =}\NormalTok{ pop}\SpecialCharTok{/}\NormalTok{area,}
         \CommentTok{\# convert region to factors and recoded them accordingly }
         \AttributeTok{region =} \FunctionTok{factor}\NormalTok{(region, }\AttributeTok{levels =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{,}
                    \AttributeTok{labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"northeast"}\NormalTok{, }\StringTok{"northcentral"}\NormalTok{, }\StringTok{"south"}\NormalTok{, }\StringTok{"west"}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(docs,beds))}
\end{Highlighting}
\end{Shaded}

\hypertarget{descriptive-statistics-of-all-varaibles}{%
\subsection{Descriptive statistics of all
varaibles}\label{descriptive-statistics-of-all-varaibles}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cdi\_descriptive }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(id,cty,state,region))}

\CommentTok{\# Global}
\NormalTok{skimr}\SpecialCharTok{::}\FunctionTok{skim}\NormalTok{(cdi\_descriptive) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(}\StringTok{"skim\_type"}\NormalTok{,}\StringTok{"complete\_rate"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{skim\_variable =} 
             \FunctionTok{recode}\NormalTok{(skim\_variable, }\AttributeTok{pcincome =} \StringTok{"pcincome (in dollars)"}\NormalTok{, }\AttributeTok{totalinc =} \StringTok{"totalinc (in million of dollars)"}
\NormalTok{  )) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  knitr}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(}
    \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"variable"}\NormalTok{, }\StringTok{"n\_missing"}\NormalTok{, }\StringTok{"mean"}\NormalTok{,}\StringTok{"sd"}\NormalTok{,}\StringTok{"min"}\NormalTok{,}\StringTok{"Q25"}\NormalTok{,}\StringTok{"median"}\NormalTok{,}\StringTok{"Q75"}\NormalTok{,}\StringTok{"max"}\NormalTok{,}\StringTok{"histogram"}\NormalTok{),}
    \AttributeTok{caption =} \StringTok{"Global Summary"}\NormalTok{, }\AttributeTok{digits =} \DecValTok{4}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.24}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.07}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.09}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.09}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.09}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.09}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.09}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.09}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.09}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.07}}@{}}
\caption{Global Summary}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
n\_missing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
mean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
sd
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
min
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Q25
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
median
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Q75
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
max
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
histogram
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
n\_missing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
mean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
sd
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
min
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Q25
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
median
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Q75
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
max
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
histogram
\end{minipage} \\
\midrule
\endhead
area & 0 & 1041.4114 & 1549.9221 & 15.0000 & 451.2500 & 656.5000 &
946.7500 & 20062.0000 & ▇▁▁▁▁ \\
pop & 0 & 393010.9205 & 601987.0165 & 100043.0000 & 139027.2500 &
217280.5000 & 436064.5000 & 8863164.0000 & ▇▁▁▁▁ \\
pop18 & 0 & 28.5684 & 4.1911 & 16.4000 & 26.2000 & 28.1000 & 30.0250 &
49.7000 & ▁▇▃▁▁ \\
pop65 & 0 & 12.1698 & 3.9927 & 3.0000 & 9.8750 & 11.7500 & 13.6250 &
33.8000 & ▂▇▁▁▁ \\
crimes & 0 & 27111.6182 & 58237.5064 & 563.0000 & 6219.5000 & 11820.5000
& 26279.5000 & 688936.0000 & ▇▁▁▁▁ \\
hsgrad & 0 & 77.5607 & 7.0152 & 46.6000 & 73.8750 & 77.7000 & 82.4000 &
92.9000 & ▁▁▃▇▃ \\
bagrad & 0 & 21.0811 & 7.6545 & 8.1000 & 15.2750 & 19.7000 & 25.3250 &
52.3000 & ▆▇▃▁▁ \\
poverty & 0 & 8.7207 & 4.6567 & 1.4000 & 5.3000 & 7.9000 & 10.9000 &
36.3000 & ▇▆▁▁▁ \\
unemp & 0 & 6.5966 & 2.3379 & 2.2000 & 5.1000 & 6.2000 & 7.5000 &
21.3000 & ▇▇▁▁▁ \\
pcincome (in dollars) & 0 & 18561.4818 & 4059.1920 & 8899.0000 &
16118.2500 & 17759.0000 & 20270.0000 & 37541.0000 & ▁▇▂▁▁ \\
totalinc (in million of dollars) & 0 & 7869.2727 & 12884.3215 &
1141.0000 & 2311.0000 & 3857.0000 & 8654.2500 & 184230.0000 & ▇▁▁▁▁ \\
crm\_1000 & 0 & 57.2864 & 27.3277 & 4.6014 & 38.1019 & 52.4286 & 72.5969
& 295.9867 & ▇▅▁▁▁ \\
pdocs\_1000 & 0 & 2.1230 & 1.5329 & 0.3559 & 1.2127 & 1.7509 & 2.4915 &
17.0377 & ▇▁▁▁▁ \\
pbeds\_1000 & 0 & 3.6493 & 2.0011 & 0.1649 & 2.1972 & 3.3287 & 4.5649 &
19.6982 & ▇▃▁▁▁ \\
density\_pop & 0 & 888.4388 & 2194.7231 & 13.2587 & 192.3449 & 335.9081
& 756.5516 & 32403.7183 & ▇▁▁▁▁ \\
\bottomrule
\end{longtable}

\hypertarget{descriptive-analysis}{%
\subsection{Descriptive analysis}\label{descriptive-analysis}}

We can use the box plot/ or histogram to check for normality. But I
forgot when do we need normality\ldots isn't it for residual??

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{par}\NormalTok{(}\AttributeTok{mfrow =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{7}\NormalTok{))}

\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{density\_pop, }\AttributeTok{main =} \StringTok{"density\_pop"}\NormalTok{)}
\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{area, }\AttributeTok{main =} \StringTok{"area"}\NormalTok{)}

\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{pop, }\AttributeTok{main =} \StringTok{"pop"}\NormalTok{)}
\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{pop18, }\AttributeTok{main =} \StringTok{"pop18"}\NormalTok{)}
\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{pop65, }\AttributeTok{main =} \StringTok{"pop65"}\NormalTok{)}

\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{pdocs\_1000, }\AttributeTok{main =} \StringTok{"pdocs\_1000"}\NormalTok{)}
\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{pbeds\_1000, }\AttributeTok{main =} \StringTok{"pbeds\_1000"}\NormalTok{)}

\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{crm\_1000,  }\AttributeTok{main =} \StringTok{"crm\_1000"}\NormalTok{)}

\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{hsgrad, }\AttributeTok{main =} \StringTok{"hsgrad"}\NormalTok{)}
\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{bagrad, }\AttributeTok{main =} \StringTok{"bagrad"}\NormalTok{)}
\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{poverty, }\AttributeTok{main =} \StringTok{"poverty"}\NormalTok{)}
\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{unemp,}\AttributeTok{main =} \StringTok{"unemp"}\NormalTok{)}
\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{pcincome, }\AttributeTok{main =} \StringTok{"pcincome"}\NormalTok{)}
\FunctionTok{boxplot}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{totalinc, }\AttributeTok{main =} \StringTok{"totalinc"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/boxplot-1.pdf}

\hypertarget{correlation}{%
\subsection{Correlation}\label{correlation}}

\hypertarget{pairwise-relationship}{%
\subsubsection{Pairwise relationship}\label{pairwise-relationship}}

\begin{itemize}
\tightlist
\item
  This gives us an idea of the correlation between each variable, but my
  old project build whole model first then assessed the correlation.
  Need Discussion!
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(corrplot)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## corrplot 0.92 loaded
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{cor}\NormalTok{(cdi\_descriptive) }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ knitr}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.07}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.06}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 30\tabcolsep) * \real{0.07}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
area
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
pop
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
pop18
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
pop65
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
crimes
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
hsgrad
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
bagrad
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
poverty
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
unemp
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
pcincome
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
totalinc
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
crm\_1000
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
pdocs\_1000
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
pbeds\_1000
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
density\_pop
\end{minipage} \\
\midrule
\endhead
area & 1.0000000 & 0.1730834 & -0.0548781 & 0.0057709 & 0.1294754 &
-0.0985981 & -0.1372377 & 0.1713433 & 0.1992093 & -0.1877151 & 0.1270743
& 0.0429484 & -0.1163860 & -0.1412335 & -0.1568156 \\
pop & 0.1730834 & 1.0000000 & 0.0783721 & -0.0290374 & 0.8863318 &
-0.0174269 & 0.1468138 & 0.0380195 & 0.0053517 & 0.2356102 & 0.9867476 &
0.2800992 & 0.1668595 & 0.0203012 & 0.3220266 \\
pop18 & -0.0548781 & 0.0783721 & 1.0000000 & -0.6163096 & 0.0899406 &
0.2505843 & 0.4560970 & 0.0339755 & -0.2785271 & -0.0316484 & 0.0711615
& 0.1905688 & 0.2370280 & 0.0295244 & 0.1254644 \\
pop65 & 0.0057709 & -0.0290374 & -0.6163096 & 1.0000000 & -0.0352903 &
-0.2682518 & -0.3392288 & 0.0065785 & 0.2363094 & 0.0185907 & -0.0227332
& -0.0665333 & 0.0186087 & 0.2471479 & 0.0291845 \\
crimes & 0.1294754 & 0.8863318 & 0.0899406 & -0.0352903 & 1.0000000 &
-0.1063284 & 0.0770765 & 0.1644057 & 0.0435568 & 0.1175391 & 0.8430980 &
0.5300430 & 0.1577103 & 0.0778907 & 0.5609842 \\
hsgrad & -0.0985981 & -0.0174269 & 0.2505843 & -0.2682518 & -0.1063284 &
1.0000000 & 0.7077867 & -0.6917505 & -0.5935958 & 0.5229961 & 0.0433557
& -0.2264129 & 0.1427765 & -0.2111625 & -0.1040070 \\
bagrad & -0.1372377 & 0.1468138 & 0.4560970 & -0.3392288 & 0.0770765 &
0.7077867 & 1.0000000 & -0.4084238 & -0.5409069 & 0.6953619 & 0.2222301
& 0.0383046 & 0.4410463 & -0.0454183 & 0.1556063 \\
poverty & 0.1713433 & 0.0380195 & 0.0339755 & 0.0065785 & 0.1644057 &
-0.6917505 & -0.4084238 & 1.0000000 & 0.4369472 & -0.6017250 &
-0.0387393 & 0.4718442 & 0.0637048 & 0.3713989 & 0.1265079 \\
unemp & 0.1992093 & 0.0053517 & -0.2785271 & 0.2363094 & 0.0435568 &
-0.5935958 & -0.5409069 & 0.4369472 & 1.0000000 & -0.3221444 &
-0.0338763 & 0.0418466 & -0.2478866 & -0.0624878 & 0.0227179 \\
pcincome & -0.1877151 & 0.2356102 & -0.0316484 & 0.0185907 & 0.1175391 &
0.5229961 & 0.6953619 & -0.6017250 & -0.3221444 & 1.0000000 & 0.3476816
& -0.0802442 & 0.3600458 & -0.0535500 & 0.2332260 \\
totalinc & 0.1270743 & 0.9867476 & 0.0711615 & -0.0227332 & 0.8430980 &
0.0433557 & 0.2222301 & -0.0387393 & -0.0338763 & 0.3476816 & 1.0000000
& 0.2281557 & 0.1991038 & 0.0063239 & 0.3162048 \\
crm\_1000 & 0.0429484 & 0.2800992 & 0.1905688 & -0.0665333 & 0.5300430 &
-0.2264129 & 0.0383046 & 0.4718442 & 0.0418466 & -0.0802442 & 0.2281557
& 1.0000000 & 0.3070831 & 0.3644505 & 0.4804285 \\
pdocs\_1000 & -0.1163860 & 0.1668595 & 0.2370280 & 0.0186087 & 0.1577103
& 0.1427765 & 0.4410463 & 0.0637048 & -0.2478866 & 0.3600458 & 0.1991038
& 0.3070831 & 1.0000000 & 0.6666947 & 0.3180424 \\
pbeds\_1000 & -0.1412335 & 0.0203012 & 0.0295244 & 0.2471479 & 0.0778907
& -0.2111625 & -0.0454183 & 0.3713989 & -0.0624878 & -0.0535500 &
0.0063239 & 0.3644505 & 0.6666947 & 1.0000000 & 0.2064177 \\
density\_pop & -0.1568156 & 0.3220266 & 0.1254644 & 0.0291845 &
0.5609842 & -0.1040070 & 0.1556063 & 0.1265079 & 0.0227179 & 0.2332260 &
0.3162048 & 0.4804285 & 0.3180424 & 0.2064177 & 1.0000000 \\
\bottomrule
\end{longtable}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggcorrplot)}

\FunctionTok{ggcorrplot}\NormalTok{(}\FunctionTok{cor}\NormalTok{(cdi\_descriptive), }\AttributeTok{type =} \StringTok{"upper"}\NormalTok{,}
   \AttributeTok{lab =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-2-1.pdf}

Scatter plot

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{pairs}\NormalTok{(crm\_1000 }\SpecialCharTok{\textasciitilde{}}\NormalTok{.,}\AttributeTok{data=}\NormalTok{cdi\_descriptive, }\AttributeTok{panel =}\NormalTok{ panel.smooth, }\AttributeTok{upper.panel =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{main =} \StringTok{"Scatterplot Matrix"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-3-1.pdf}

\hypertarget{marginal-distribution}{%
\subsection{Marginal distribution}\label{marginal-distribution}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(ggExtra)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_den }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ density\_pop, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_den, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-5-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_area }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ area, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_area, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-6-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_pop }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ pop, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_pop, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-7-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_pop18 }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ pop18, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\CommentTok{\# positive correlation}
\FunctionTok{ggMarginal}\NormalTok{(marg\_pop18, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-8-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_pop65 }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ pop65, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_pop65, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-9-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_pdocs\_1000 }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ pdocs\_1000, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_pdocs\_1000, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-10-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_pbeds\_1000 }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ pbeds\_1000, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_pbeds\_1000, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-11-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_hsgrad }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ hsgrad, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{) }\CommentTok{\#negative correlation}
\FunctionTok{ggMarginal}\NormalTok{(marg\_hsgrad, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-12-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_bagrad }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ bagrad, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_bagrad, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-13-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_poverty }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ poverty, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{) }\CommentTok{\# positive correlation}
\FunctionTok{ggMarginal}\NormalTok{(marg\_poverty, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-14-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_unemp }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ unemp, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_unemp, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-15-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_pcincome }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ pcincome, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_pcincome, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-16-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{marg\_totalinc }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ totalinc, }\AttributeTok{y =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{\textquotesingle{}lm\textquotesingle{}}\NormalTok{, }\AttributeTok{se =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}red\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggMarginal}\NormalTok{(marg\_totalinc, }\AttributeTok{type =} \StringTok{"histogram"}\NormalTok{, }\AttributeTok{fill=}\StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-17-1.pdf}

\hypertarget{distribution-of-outcome}{%
\subsection{Distribution of outcome}\label{distribution-of-outcome}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cdi }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ crm\_1000)) }\SpecialCharTok{+}
  \FunctionTok{geom\_histogram}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
\end{verbatim}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-18-1.pdf}

do we look at the distribution of outcome like this and transform them
here? check again

\hypertarget{counties-with-unusual-rates}{%
\subsection{Counties with unusual
rates}\label{counties-with-unusual-rates}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{upper }\OtherTok{=} \FunctionTok{quantile}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{crm\_1000, }\FloatTok{0.75}\NormalTok{)}
\NormalTok{lower }\OtherTok{=} \FunctionTok{quantile}\NormalTok{(cdi}\SpecialCharTok{$}\NormalTok{crm\_1000, }\FloatTok{0.25}\NormalTok{)}
\NormalTok{IQR }\OtherTok{=}\NormalTok{ upper }\SpecialCharTok{{-}}\NormalTok{ lower}
\NormalTok{cdi }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(crm\_1000 }\SpecialCharTok{\textgreater{}}\NormalTok{ upper }\SpecialCharTok{+} \FloatTok{1.5}\SpecialCharTok{*}\NormalTok{IQR,}
\NormalTok{         crm\_1000 }\SpecialCharTok{\textgreater{}}\NormalTok{ lower }\SpecialCharTok{{-}} \FloatTok{1.5}\SpecialCharTok{*}\NormalTok{IQR) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(cty, crm\_1000) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  knitr}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lr@{}}
\toprule
cty & crm\_1000 \\
\midrule
\endhead
Kings & 295.99 \\
Dade & 126.34 \\
Fulton & 143.35 \\
St.\_Loui & 161.60 \\
\bottomrule
\end{longtable}

\hypertarget{model}{%
\section{Model}\label{model}}

\hypertarget{full-model-predictors}{%
\subsection{Full model predictors}\label{full-model-predictors}}

this model used \texttt{northeast} as the reference level for region

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cdi\_model }\OtherTok{=}\NormalTok{ cdi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(id,cty,state,area,crimes,totalinc))}

\CommentTok{\# use }
\NormalTok{full\_fit }\OtherTok{=} \FunctionTok{lm}\NormalTok{(crm\_1000 }\SpecialCharTok{\textasciitilde{}}\NormalTok{ ., }\AttributeTok{data =}\NormalTok{ cdi\_model)}
\FunctionTok{summary}\NormalTok{(full\_fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = crm_1000 ~ ., data = cdi_model)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -47.786 -11.422  -0.934  10.200  75.180 
## 
## Coefficients:
##                      Estimate Std. Error t value Pr(>|t|)    
## (Intercept)        -6.922e+01  2.739e+01  -2.528 0.011849 *  
## pop                 5.486e-06  1.579e-06   3.474 0.000566 ***
## pop18               6.947e-01  3.305e-01   2.102 0.036150 *  
## pop65              -1.998e-01  3.055e-01  -0.654 0.513410    
## hsgrad              6.143e-01  2.690e-01   2.284 0.022864 *  
## bagrad             -4.835e-01  2.971e-01  -1.628 0.104327    
## poverty             1.856e+00  3.864e-01   4.803 2.17e-06 ***
## unemp               6.111e-01  5.314e-01   1.150 0.250812    
## pcincome            1.039e-03  4.734e-04   2.195 0.028670 *  
## regionnorthcentral  8.978e+00  2.732e+00   3.286 0.001100 ** 
## regionsouth         2.779e+01  2.659e+00  10.453  < 2e-16 ***
## regionwest          2.118e+01  3.125e+00   6.778 4.09e-11 ***
## pdocs_1000         -6.634e-01  1.019e+00  -0.651 0.515556    
## pbeds_1000          3.157e+00  7.939e-01   3.977 8.21e-05 ***
## density_pop         4.901e-03  4.537e-04  10.802  < 2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 17.81 on 425 degrees of freedom
## Multiple R-squared:  0.589,  Adjusted R-squared:  0.5755 
## F-statistic: 43.51 on 14 and 425 DF,  p-value: < 2.2e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_qq}\NormalTok{(full\_fit)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-20-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_fit}\NormalTok{(full\_fit)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-20-2.pdf}

\hypertarget{transformation}{%
\subsection{Transformation}\label{transformation}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{lambda }\OtherTok{=}\NormalTok{ MASS}\SpecialCharTok{::}\FunctionTok{boxcox}\NormalTok{(full\_fit)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-21-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{optlam }\OtherTok{=}\NormalTok{ lambda}\SpecialCharTok{$}\NormalTok{x[}\FunctionTok{which.max}\NormalTok{(lambda}\SpecialCharTok{$}\NormalTok{y)]}
\NormalTok{optlam}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 0.5454545
\end{verbatim}

The lambda from the transformation is 0.5454, so we will try to fit a
square root transformation to Y

New tibble with transformed Y:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cdi\_trans }\OtherTok{=}\NormalTok{ cdi\_model }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{crm\_1000\_sqr =}\NormalTok{ crm\_1000}\SpecialCharTok{\^{}}\FloatTok{0.5}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(crm\_1000))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{trans\_fit }\OtherTok{=} \FunctionTok{lm}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{ .,}\AttributeTok{data =}\NormalTok{ cdi\_trans)}
\FunctionTok{summary}\NormalTok{(trans\_fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = crm_1000_sqr ~ ., data = cdi_trans)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -4.0410 -0.7300  0.0708  0.7485  4.0273 
## 
## Coefficients:
##                      Estimate Std. Error t value Pr(>|t|)    
## (Intercept)        -1.645e+00  1.802e+00  -0.913 0.361766    
## pop                 3.624e-07  1.039e-07   3.488 0.000537 ***
## pop18               6.320e-02  2.174e-02   2.907 0.003844 ** 
## pop65              -3.609e-03  2.009e-02  -0.180 0.857548    
## hsgrad              3.479e-02  1.769e-02   1.966 0.049933 *  
## bagrad             -3.472e-02  1.954e-02  -1.777 0.076307 .  
## poverty             1.192e-01  2.542e-02   4.688 3.72e-06 ***
## unemp               4.305e-02  3.496e-02   1.232 0.218783    
## pcincome            9.589e-05  3.114e-05   3.079 0.002213 ** 
## regionnorthcentral  7.062e-01  1.797e-01   3.929 9.94e-05 ***
## regionsouth         1.996e+00  1.749e-01  11.413  < 2e-16 ***
## regionwest          1.674e+00  2.056e-01   8.142 4.37e-15 ***
## pdocs_1000         -3.845e-02  6.706e-02  -0.573 0.566744    
## pbeds_1000          2.120e-01  5.223e-02   4.059 5.86e-05 ***
## density_pop         2.150e-04  2.984e-05   7.203 2.69e-12 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 1.171 on 425 degrees of freedom
## Multiple R-squared:  0.5602, Adjusted R-squared:  0.5457 
## F-statistic: 38.66 on 14 and 425 DF,  p-value: < 2.2e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_fit}\NormalTok{(trans\_fit)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-23-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_qq}\NormalTok{(trans\_fit)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-23-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{lambda\_trans }\OtherTok{=}\NormalTok{ MASS}\SpecialCharTok{::}\FunctionTok{boxcox}\NormalTok{(trans\_fit)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-23-3.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{optlam\_trans }\OtherTok{=}\NormalTok{ lambda\_trans}\SpecialCharTok{$}\NormalTok{x[}\FunctionTok{which.max}\NormalTok{(lambda\_trans}\SpecialCharTok{$}\NormalTok{y)]}
\NormalTok{optlam\_trans}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.111111
\end{verbatim}

\hypertarget{counties-with-unusual-rates-1}{%
\subsubsection{Counties with unusual
rates}\label{counties-with-unusual-rates-1}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{upper }\OtherTok{=} \FunctionTok{quantile}\NormalTok{(cdi\_trans}\SpecialCharTok{$}\NormalTok{crm\_1000\_sqr, }\FloatTok{0.75}\NormalTok{)}
\NormalTok{lower }\OtherTok{=} \FunctionTok{quantile}\NormalTok{(cdi\_trans}\SpecialCharTok{$}\NormalTok{crm\_1000\_sqr, }\FloatTok{0.25}\NormalTok{)}
\NormalTok{IQR }\OtherTok{=}\NormalTok{ upper }\SpecialCharTok{{-}}\NormalTok{ lower}

\NormalTok{cdi\_trans }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textgreater{}}\NormalTok{ upper }\SpecialCharTok{+} \FloatTok{1.5}\SpecialCharTok{*}\NormalTok{IQR,}
\NormalTok{         crm\_1000\_sqr }\SpecialCharTok{\textgreater{}}\NormalTok{ lower }\SpecialCharTok{{-}} \FloatTok{1.5}\SpecialCharTok{*}\NormalTok{IQR) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(crm\_1000\_sqr) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cty =} \FunctionTok{c}\NormalTok{(}\StringTok{"Kings"}\NormalTok{, }\StringTok{"St.\_Loui"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  knitr}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(}\AttributeTok{digits =} \DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rl@{}}
\toprule
crm\_1000\_sqr & cty \\
\midrule
\endhead
17.20 & Kings \\
12.71 & St.\_Loui \\
\bottomrule
\end{longtable}

\hypertarget{model-selection}{%
\subsection{Model Selection}\label{model-selection}}

\hypertarget{step-wise-approach}{%
\subsubsection{Step-wise approach}\label{step-wise-approach}}

\hypertarget{backward}{%
\paragraph{Backward}\label{backward}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_back }\OtherTok{=} \FunctionTok{step}\NormalTok{(trans\_fit, }\AttributeTok{direction=}\StringTok{\textquotesingle{}backward\textquotesingle{}}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Start:  AIC=153.85
## crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + poverty + 
##     unemp + pcincome + region + pdocs_1000 + pbeds_1000 + density_pop
## 
##               Df Sum of Sq    RSS    AIC
## - pop65        1     0.044 583.09 151.89
## - pdocs_1000   1     0.451 583.49 152.19
## - unemp        1     2.081 585.12 153.42
## <none>                     583.04 153.85
## - bagrad       1     4.331 587.37 155.11
## - hsgrad       1     5.303 588.35 155.84
## - pop18        1    11.591 594.63 160.51
## - pcincome     1    13.004 596.05 161.56
## - pop          1    16.690 599.73 164.27
## - pbeds_1000   1    22.604 605.65 168.59
## - poverty      1    30.154 613.20 174.04
## - density_pop  1    71.185 654.23 202.54
## - region       3   205.212 788.25 280.54
## 
## Step:  AIC=151.89
## crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + 
##     pcincome + region + pdocs_1000 + pbeds_1000 + density_pop
## 
##               Df Sum of Sq    RSS    AIC
## - pdocs_1000   1     0.446 583.53 150.22
## - unemp        1     2.038 585.12 151.42
## <none>                     583.09 151.89
## - bagrad       1     4.334 587.42 153.15
## - hsgrad       1     5.380 588.47 153.93
## - pcincome     1    13.159 596.25 159.71
## - pop18        1    16.404 599.49 162.09
## - pop          1    16.651 599.74 162.28
## - pbeds_1000   1    23.989 607.08 167.63
## - poverty      1    31.904 614.99 173.33
## - density_pop  1    71.276 654.36 200.63
## - region       3   205.335 788.42 278.63
## 
## Step:  AIC=150.22
## crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + 
##     pcincome + region + pbeds_1000 + density_pop
## 
##               Df Sum of Sq    RSS    AIC
## - unemp        1     2.006 585.54 149.73
## <none>                     583.53 150.22
## - bagrad       1     5.378 588.91 152.26
## - hsgrad       1     5.671 589.20 152.48
## - pcincome     1    12.753 596.29 157.74
## - pop18        1    16.012 599.54 160.13
## - pop          1    16.517 600.05 160.50
## - poverty      1    32.264 615.80 171.90
## - pbeds_1000   1    41.083 624.62 178.16
## - density_pop  1    70.863 654.40 198.65
## - region       3   204.905 788.44 276.64
## 
## Step:  AIC=149.73
## crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + 
##     region + pbeds_1000 + density_pop
## 
##               Df Sum of Sq    RSS    AIC
## <none>                     585.54 149.73
## - hsgrad       1     4.640 590.18 151.21
## - bagrad       1     7.070 592.61 153.01
## - pcincome     1    15.076 600.61 158.92
## - pop18        1    15.967 601.50 159.57
## - pop          1    16.046 601.58 159.63
## - pbeds_1000   1    39.273 624.81 176.30
## - poverty      1    41.161 626.70 177.62
## - density_pop  1    70.118 655.66 197.50
## - region       3   209.818 795.36 278.49
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_back}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + 
##     pcincome + region + pbeds_1000 + density_pop, data = cdi_trans)
## 
## Coefficients:
##        (Intercept)                 pop               pop18              hsgrad  
##         -1.132e+00           3.546e-07           6.390e-02           3.180e-02  
##             bagrad             poverty            pcincome  regionnorthcentral  
##         -4.210e-02           1.298e-01           1.005e-04           6.887e-01  
##        regionsouth          regionwest          pbeds_1000         density_pop  
##          1.933e+00           1.642e+00           1.767e-01           2.117e-04
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_fit}\NormalTok{(fit\_back)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-25-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_qq}\NormalTok{(fit\_back)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-25-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{MASS}\SpecialCharTok{::}\FunctionTok{boxcox}\NormalTok{(fit\_back)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-25-3.pdf}

Backward: crm\_1000\_sqr \textasciitilde{} pop + pop18 + hsgrad + bagrad
+ poverty + pcincome + region + pbeds\_1000 + density\_pop

\hypertarget{forward}{%
\paragraph{forward}\label{forward}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_forward }\OtherTok{=} \FunctionTok{step}\NormalTok{(trans\_fit, }\AttributeTok{direction=}\StringTok{"forward"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Start:  AIC=153.85
## crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + poverty + 
##     unemp + pcincome + region + pdocs_1000 + pbeds_1000 + density_pop
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_forward}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + 
##     poverty + unemp + pcincome + region + pdocs_1000 + pbeds_1000 + 
##     density_pop, data = cdi_trans)
## 
## Coefficients:
##        (Intercept)                 pop               pop18               pop65  
##         -1.645e+00           3.624e-07           6.320e-02          -3.609e-03  
##             hsgrad              bagrad             poverty               unemp  
##          3.479e-02          -3.472e-02           1.192e-01           4.305e-02  
##           pcincome  regionnorthcentral         regionsouth          regionwest  
##          9.589e-05           7.062e-01           1.996e+00           1.674e+00  
##         pdocs_1000          pbeds_1000         density_pop  
##         -3.845e-02           2.120e-01           2.150e-04
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_fit}\NormalTok{(fit\_forward)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-26-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_qq}\NormalTok{(fit\_forward)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-26-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{MASS}\SpecialCharTok{::}\FunctionTok{boxcox}\NormalTok{(fit\_forward)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-26-3.pdf}

model: crm\_1000\_sqr \textasciitilde{} pop + pop18 + pop65 + hsgrad +
bagrad + poverty + unemp + pcincome + region + pdocs\_1000 + pbeds\_1000
+ density\_pop, data = cdi\_trans)

\hypertarget{both}{%
\paragraph{both}\label{both}}

step-wise

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_both }\OtherTok{=} \FunctionTok{step}\NormalTok{(trans\_fit, }\AttributeTok{direction=}\StringTok{\textquotesingle{}both\textquotesingle{}}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Start:  AIC=153.85
## crm_1000_sqr ~ pop + pop18 + pop65 + hsgrad + bagrad + poverty + 
##     unemp + pcincome + region + pdocs_1000 + pbeds_1000 + density_pop
## 
##               Df Sum of Sq    RSS    AIC
## - pop65        1     0.044 583.09 151.89
## - pdocs_1000   1     0.451 583.49 152.19
## - unemp        1     2.081 585.12 153.42
## <none>                     583.04 153.85
## - bagrad       1     4.331 587.37 155.11
## - hsgrad       1     5.303 588.35 155.84
## - pop18        1    11.591 594.63 160.51
## - pcincome     1    13.004 596.05 161.56
## - pop          1    16.690 599.73 164.27
## - pbeds_1000   1    22.604 605.65 168.59
## - poverty      1    30.154 613.20 174.04
## - density_pop  1    71.185 654.23 202.54
## - region       3   205.212 788.25 280.54
## 
## Step:  AIC=151.89
## crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + 
##     pcincome + region + pdocs_1000 + pbeds_1000 + density_pop
## 
##               Df Sum of Sq    RSS    AIC
## - pdocs_1000   1     0.446 583.53 150.22
## - unemp        1     2.038 585.12 151.42
## <none>                     583.09 151.89
## - bagrad       1     4.334 587.42 153.15
## + pop65        1     0.044 583.04 153.85
## - hsgrad       1     5.380 588.47 153.93
## - pcincome     1    13.159 596.25 159.71
## - pop18        1    16.404 599.49 162.09
## - pop          1    16.651 599.74 162.28
## - pbeds_1000   1    23.989 607.08 167.63
## - poverty      1    31.904 614.99 173.33
## - density_pop  1    71.276 654.36 200.63
## - region       3   205.335 788.42 278.63
## 
## Step:  AIC=150.22
## crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + 
##     pcincome + region + pbeds_1000 + density_pop
## 
##               Df Sum of Sq    RSS    AIC
## - unemp        1     2.006 585.54 149.73
## <none>                     583.53 150.22
## + pdocs_1000   1     0.446 583.09 151.89
## + pop65        1     0.039 583.49 152.19
## - bagrad       1     5.378 588.91 152.26
## - hsgrad       1     5.671 589.20 152.48
## - pcincome     1    12.753 596.29 157.74
## - pop18        1    16.012 599.54 160.13
## - pop          1    16.517 600.05 160.50
## - poverty      1    32.264 615.80 171.90
## - pbeds_1000   1    41.083 624.62 178.16
## - density_pop  1    70.863 654.40 198.65
## - region       3   204.905 788.44 276.64
## 
## Step:  AIC=149.73
## crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + 
##     region + pbeds_1000 + density_pop
## 
##               Df Sum of Sq    RSS    AIC
## <none>                     585.54 149.73
## + unemp        1     2.006 583.53 150.22
## - hsgrad       1     4.640 590.18 151.21
## + pdocs_1000   1     0.414 585.12 151.42
## + pop65        1     0.000 585.54 151.73
## - bagrad       1     7.070 592.61 153.01
## - pcincome     1    15.076 600.61 158.92
## - pop18        1    15.967 601.50 159.57
## - pop          1    16.046 601.58 159.63
## - pbeds_1000   1    39.273 624.81 176.30
## - poverty      1    41.161 626.70 177.62
## - density_pop  1    70.118 655.66 197.50
## - region       3   209.818 795.36 278.49
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_both}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + 
##     pcincome + region + pbeds_1000 + density_pop, data = cdi_trans)
## 
## Coefficients:
##        (Intercept)                 pop               pop18              hsgrad  
##         -1.132e+00           3.546e-07           6.390e-02           3.180e-02  
##             bagrad             poverty            pcincome  regionnorthcentral  
##         -4.210e-02           1.298e-01           1.005e-04           6.887e-01  
##        regionsouth          regionwest          pbeds_1000         density_pop  
##          1.933e+00           1.642e+00           1.767e-01           2.117e-04
\end{verbatim}

lm(formula = crm\_1000\^{}0.5 \textasciitilde{} pop + pop18 + hsgrad +
bagrad + poverty + pcincome + region + pbeds\_1000 + density\_pop, data
= cdi\_model)

This gives back the same model as back

\hypertarget{criterion-based-adjr2-and-cp}{%
\subsubsection{Criterion Based: Adjr2 and
Cp}\label{criterion-based-adjr2-and-cp}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# printing the 2 best models of each size. For example, the first two lines: print the best 2 models that have 2 variables (including intercept)}

\NormalTok{models }\OtherTok{\textless{}{-}} \FunctionTok{regsubsets}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{., }\AttributeTok{data =}\NormalTok{ cdi\_trans,}\AttributeTok{nvmax =} \DecValTok{12}\NormalTok{)}
\NormalTok{res\_sum }\OtherTok{=} \FunctionTok{summary}\NormalTok{(models)}

\FunctionTok{par}\NormalTok{(}\AttributeTok{mfrow=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{))}
\FunctionTok{plot}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{12}\NormalTok{, res\_sum}\SpecialCharTok{$}\NormalTok{cp, }\AttributeTok{xlab=}\StringTok{"No of parameters"}\NormalTok{, }\AttributeTok{ylab=}\StringTok{"Cp Statistic"}\NormalTok{)}
\FunctionTok{plot}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{12}\NormalTok{, res\_sum}\SpecialCharTok{$}\NormalTok{adjr2, }\AttributeTok{xlab=}\StringTok{"No of parameters"}\NormalTok{, }\AttributeTok{ylab=}\StringTok{"Adj R2"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-28-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mod\_s }\OtherTok{=} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{Adj.R2 =} \FunctionTok{which.max}\NormalTok{(res\_sum}\SpecialCharTok{$}\NormalTok{adjr2),}
  \AttributeTok{CP =} \FunctionTok{which.min}\NormalTok{(res\_sum}\SpecialCharTok{$}\NormalTok{cp)}
\NormalTok{)}

\NormalTok{get\_model\_formula }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(id, object, outcome)\{}
  \CommentTok{\# get models data}
\NormalTok{  models }\OtherTok{\textless{}{-}} \FunctionTok{summary}\NormalTok{(object)}\SpecialCharTok{$}\NormalTok{which[id,}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{]}
  \CommentTok{\# Get outcome variable}
  \CommentTok{\#form \textless{}{-} as.formula(object$call[[2]])}
  \CommentTok{\#outcome \textless{}{-} all.vars(form)[1]}
  \CommentTok{\# Get model predictors}
\NormalTok{  predictors }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(}\FunctionTok{which}\NormalTok{(models }\SpecialCharTok{==} \ConstantTok{TRUE}\NormalTok{))}
\NormalTok{  predictors }\OtherTok{\textless{}{-}} \FunctionTok{paste}\NormalTok{(predictors, }\AttributeTok{collapse =} \StringTok{"+"}\NormalTok{)}
  \CommentTok{\# Build model formula}
  \FunctionTok{as.formula}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(outcome, }\StringTok{"\textasciitilde{}"}\NormalTok{, predictors))}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Model chosen by Adj \(R^2\):

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{get\_model\_formula}\NormalTok{(}\DecValTok{12}\NormalTok{, models, }\StringTok{"crm\_1000\_sqr"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + 
##     pcincome + regionnorthcentral + regionsouth + regionwest + 
##     pbeds_1000 + density_pop
## <environment: 0x7fed56451768>
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_adjr }\OtherTok{=} \FunctionTok{lm}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{ pop }\SpecialCharTok{+}\NormalTok{ pop18 }\SpecialCharTok{+}\NormalTok{ hsgrad }\SpecialCharTok{+}\NormalTok{ bagrad }\SpecialCharTok{+}\NormalTok{ poverty }\SpecialCharTok{+}\NormalTok{ unemp }\SpecialCharTok{+}\NormalTok{ pcincome }\SpecialCharTok{+}\NormalTok{ region}\SpecialCharTok{+}\NormalTok{ pbeds\_1000 }\SpecialCharTok{+}\NormalTok{ density\_pop, }\AttributeTok{data =}\NormalTok{ cdi\_trans)}

\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_fit}\NormalTok{(fit\_adjr)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-30-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_qq}\NormalTok{(fit\_adjr)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-30-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{MASS}\SpecialCharTok{::}\FunctionTok{boxcox}\NormalTok{(fit\_adjr)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-30-3.pdf}

crm\_1000\_sqr \textasciitilde{} pop + pop18 + hsgrad + bagrad + poverty
+ unemp + pcincome + region+ pbeds\_1000 + density\_pop

Model chosen by Cp: Same with backward

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{get\_model\_formula}\NormalTok{(}\DecValTok{11}\NormalTok{, models, }\StringTok{"crm\_1000\_sqr"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + 
##     regionnorthcentral + regionsouth + regionwest + pbeds_1000 + 
##     density_pop
## <environment: 0x7fed562f93b0>
\end{verbatim}

crm\_1000\_sqr \textasciitilde{} pop + pop18 + hsgrad + bagrad + poverty
+ pcincome + region+ pbeds\_1000 + density\_pop

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_Cp }\OtherTok{=} \FunctionTok{lm}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{ pop }\SpecialCharTok{+}\NormalTok{ pop18 }\SpecialCharTok{+}\NormalTok{ hsgrad }\SpecialCharTok{+}\NormalTok{ bagrad }\SpecialCharTok{+}\NormalTok{ poverty }\SpecialCharTok{+}\NormalTok{ pcincome }\SpecialCharTok{+}\NormalTok{ region}\SpecialCharTok{+}\NormalTok{ pbeds\_1000 }\SpecialCharTok{+}\NormalTok{ density\_pop, }\AttributeTok{data=}\NormalTok{cdi\_trans)}

\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_fit}\NormalTok{(fit\_Cp)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-32-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_qq}\NormalTok{(fit\_Cp)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-32-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{MASS}\SpecialCharTok{::}\FunctionTok{boxcox}\NormalTok{(fit\_Cp)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-32-3.pdf}

\hypertarget{check-collinearity}{%
\subsection{Check Collinearity}\label{check-collinearity}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{performance}\SpecialCharTok{::}\FunctionTok{check\_collinearity}\NormalTok{(fit\_back)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Check for Multicollinearity
## 
## Low Correlation
## 
##         Term  VIF Increased SE Tolerance
##          pop 1.00         1.00      1.00
##        pop18 1.93         1.39      0.52
##       hsgrad 3.15         1.78      0.32
##       bagrad 3.35         1.83      0.30
##      poverty 2.16         1.47      0.46
##     pcincome 1.03         1.01      0.97
##       region 1.63         1.28      0.61
##   pbeds_1000 1.32         1.15      0.76
##  density_pop 1.01         1.00      0.99
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{performance}\SpecialCharTok{::}\FunctionTok{check\_collinearity}\NormalTok{(fit\_forward)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Check for Multicollinearity
## 
## Low Correlation
## 
##         Term  VIF Increased SE Tolerance
##          pop 1.00         1.00      1.00
##        pop18 2.60         1.61      0.38
##        pop65 2.05         1.43      0.49
##       hsgrad 3.30         1.82      0.30
##       bagrad 3.72         1.93      0.27
##      poverty 2.48         1.58      0.40
##        unemp 1.89         1.37      0.53
##     pcincome 1.02         1.01      0.98
##       region 1.99         1.41      0.50
##   pdocs_1000 2.75         1.66      0.36
##   pbeds_1000 3.31         1.82      0.30
##  density_pop 1.00         1.00      1.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{performance}\SpecialCharTok{::}\FunctionTok{check\_collinearity}\NormalTok{(fit\_adjr)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Check for Multicollinearity
## 
## Low Correlation
## 
##         Term  VIF Increased SE Tolerance
##          pop 1.00         1.00      1.00
##        pop18 1.93         1.39      0.52
##       hsgrad 3.27         1.81      0.31
##       bagrad 3.49         1.87      0.29
##      poverty 2.38         1.54      0.42
##        unemp 1.86         1.36      0.54
##     pcincome 1.03         1.01      0.97
##       region 1.83         1.35      0.55
##   pbeds_1000 1.44         1.20      0.69
##  density_pop 1.00         1.00      1.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{performance}\SpecialCharTok{::}\FunctionTok{check\_collinearity}\NormalTok{(fit\_Cp)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Check for Multicollinearity
## 
## Low Correlation
## 
##         Term  VIF Increased SE Tolerance
##          pop 1.00         1.00      1.00
##        pop18 1.93         1.39      0.52
##       hsgrad 3.15         1.78      0.32
##       bagrad 3.35         1.83      0.30
##      poverty 2.16         1.47      0.46
##     pcincome 1.03         1.01      0.97
##       region 1.63         1.28      0.61
##   pbeds_1000 1.32         1.15      0.76
##  density_pop 1.01         1.00      0.99
\end{verbatim}

Delete predictors when vif larger than 10. Non of them are greater than
10. So all of the models are good, don't have to worry about
collinearity.

\hypertarget{checking-to-outliers-and-influential-points}{%
\subsection{Checking to Outliers and Influential
Points}\label{checking-to-outliers-and-influential-points}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_cooksd\_bar}\NormalTok{(fit\_back, }\AttributeTok{print\_plot =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-34-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#remove observations whose cook\textquotesingle{}s D \textgreater{} 0.5}
\NormalTok{cdi\_trans\_without }\OtherTok{=}\NormalTok{ cdi\_trans[}\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{6}\NormalTok{),]}
\end{Highlighting}
\end{Shaded}

\hypertarget{build-model-without-outlier}{%
\subsection{Build model without
outlier}\label{build-model-without-outlier}}

Built the without model with \textbf{fit\_back}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_back\_without }\OtherTok{=} \FunctionTok{lm}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{ pop }\SpecialCharTok{+}\NormalTok{ pop18 }\SpecialCharTok{+}\NormalTok{ hsgrad }\SpecialCharTok{+}\NormalTok{ bagrad }\SpecialCharTok{+}\NormalTok{ poverty }\SpecialCharTok{+}\NormalTok{ pcincome }\SpecialCharTok{+}\NormalTok{ region }\SpecialCharTok{+}\NormalTok{ pbeds\_1000 }\SpecialCharTok{+}\NormalTok{ density\_pop, }\AttributeTok{data =}\NormalTok{ cdi\_trans\_without)}

\FunctionTok{summary}\NormalTok{(fit\_back\_without)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + 
##     pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -4.1228 -0.6365  0.0530  0.7450  3.9171 
## 
## Coefficients:
##                      Estimate Std. Error t value Pr(>|t|)    
## (Intercept)        -9.873e-01  1.576e+00  -0.626 0.531435    
## pop                 7.169e-07  1.419e-07   5.053 6.46e-07 ***
## pop18               7.526e-02  1.855e-02   4.056 5.93e-05 ***
## hsgrad              2.185e-02  1.695e-02   1.289 0.198159    
## bagrad             -4.136e-02  1.808e-02  -2.288 0.022652 *  
## poverty             1.219e-01  2.319e-02   5.254 2.35e-07 ***
## pcincome            1.121e-04  3.068e-05   3.655 0.000289 ***
## regionnorthcentral  7.058e-01  1.702e-01   4.146 4.08e-05 ***
## regionsouth         1.951e+00  1.617e-01  12.066  < 2e-16 ***
## regionwest          1.691e+00  1.971e-01   8.579  < 2e-16 ***
## pbeds_1000          2.036e-01  3.274e-02   6.217 1.21e-09 ***
## density_pop         7.859e-05  4.288e-05   1.833 0.067561 .  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 1.14 on 426 degrees of freedom
## Multiple R-squared:  0.5488, Adjusted R-squared:  0.5372 
## F-statistic: 47.11 on 11 and 426 DF,  p-value: < 2.2e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Residual performance}
\FunctionTok{par}\NormalTok{(}\AttributeTok{mfrow=}\FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{))}
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_fit}\NormalTok{(fit\_back\_without)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-35-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_qq}\NormalTok{(fit\_back\_without)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-35-2.pdf}

Build the without model with \textbf{trans\_fit} (full model)

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{trans\_fit\_without }\OtherTok{=} \FunctionTok{lm}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{., }\AttributeTok{data =}\NormalTok{ cdi\_trans\_without)}

\FunctionTok{summary}\NormalTok{(trans\_fit\_without)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = crm_1000_sqr ~ ., data = cdi_trans_without)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -4.0654 -0.6625  0.0540  0.7183  3.9085 
## 
## Coefficients:
##                      Estimate Std. Error t value Pr(>|t|)    
## (Intercept)        -1.643e+00  1.767e+00  -0.930 0.353040    
## pop                 7.281e-07  1.425e-07   5.111 4.87e-07 ***
## pop18               7.584e-02  2.159e-02   3.513 0.000491 ***
## pop65              -2.316e-04  1.965e-02  -0.012 0.990601    
## hsgrad              2.583e-02  1.733e-02   1.491 0.136820    
## bagrad             -3.462e-02  1.911e-02  -1.812 0.070658 .  
## poverty             1.111e-01  2.492e-02   4.457 1.07e-05 ***
## unemp               4.736e-02  3.407e-02   1.390 0.165214    
## pcincome            1.058e-04  3.141e-05   3.367 0.000828 ***
## regionnorthcentral  7.343e-01  1.755e-01   4.184 3.49e-05 ***
## regionsouth         2.024e+00  1.707e-01  11.857  < 2e-16 ***
## regionwest          1.719e+00  2.008e-01   8.565  < 2e-16 ***
## pdocs_1000         -2.102e-02  6.581e-02  -0.319 0.749576    
## pbeds_1000          2.286e-01  5.101e-02   4.481 9.59e-06 ***
## density_pop         8.083e-05  4.359e-05   1.854 0.064417 .  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 1.141 on 423 degrees of freedom
## Multiple R-squared:  0.551,  Adjusted R-squared:  0.5361 
## F-statistic: 37.08 on 14 and 423 DF,  p-value: < 2.2e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Residual performance}
\FunctionTok{par}\NormalTok{(}\AttributeTok{mfrow=}\FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{))}
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_fit}\NormalTok{(trans\_fit\_without)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-36-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_qq}\NormalTok{(trans\_fit\_without)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-36-2.pdf}

Built the without model with \textbf{fit\_adjr}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fit\_adjr\_without }\OtherTok{=} \FunctionTok{lm}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{ pop }\SpecialCharTok{+}\NormalTok{ pop18 }\SpecialCharTok{+}\NormalTok{ hsgrad }\SpecialCharTok{+}\NormalTok{ bagrad }\SpecialCharTok{+}\NormalTok{ poverty }\SpecialCharTok{+}\NormalTok{ unemp }\SpecialCharTok{+}\NormalTok{ pcincome }\SpecialCharTok{+}\NormalTok{ region}\SpecialCharTok{+}\NormalTok{ pbeds\_1000 }\SpecialCharTok{+}\NormalTok{ density\_pop, }\AttributeTok{data =}\NormalTok{ cdi\_trans\_without)}

\FunctionTok{summary}\NormalTok{(fit\_adjr\_without)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + 
##     unemp + pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -4.0662 -0.6619  0.0502  0.7174  3.9254 
## 
## Coefficients:
##                      Estimate Std. Error t value Pr(>|t|)    
## (Intercept)        -1.620e+00  1.638e+00  -0.989 0.323424    
## pop                 7.261e-07  1.419e-07   5.118 4.69e-07 ***
## pop18               7.546e-02  1.853e-02   4.072 5.57e-05 ***
## hsgrad              2.624e-02  1.722e-02   1.524 0.128270    
## bagrad             -3.617e-02  1.844e-02  -1.962 0.050439 .  
## poverty             1.115e-01  2.432e-02   4.584 6.01e-06 ***
## unemp               4.714e-02  3.372e-02   1.398 0.162867    
## pcincome            1.048e-04  3.108e-05   3.373 0.000811 ***
## regionnorthcentral  7.374e-01  1.715e-01   4.299 2.13e-05 ***
## regionsouth         2.025e+00  1.700e-01  11.913  < 2e-16 ***
## regionwest          1.711e+00  1.974e-01   8.667  < 2e-16 ***
## pbeds_1000          2.172e-01  3.414e-02   6.363 5.12e-10 ***
## density_pop         7.881e-05  4.283e-05   1.840 0.066502 .  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 1.138 on 425 degrees of freedom
## Multiple R-squared:  0.5509, Adjusted R-squared:  0.5382 
## F-statistic: 43.45 on 12 and 425 DF,  p-value: < 2.2e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Residual performance}
\FunctionTok{par}\NormalTok{(}\AttributeTok{mfrow=}\FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{))}
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_fit}\NormalTok{(fit\_adjr\_without)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-37-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_plot\_resid\_qq}\NormalTok{(fit\_adjr\_without)}
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-37-2.pdf}

\hypertarget{criteria}{%
\subsection{Criteria}\label{criteria}}

\hypertarget{cp-for-models-fit-without-outlier}{%
\paragraph{Cp for models fit without
outlier}\label{cp-for-models-fit-without-outlier}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_mallows\_cp}\NormalTok{(fit\_back\_without,trans\_fit) }\CommentTok{\#{-}14.65}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -14.65265
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_mallows\_cp}\NormalTok{(trans\_fit\_without,trans\_fit)}\CommentTok{\# {-}10.59}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -10.5956
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{olsrr}\SpecialCharTok{::}\FunctionTok{ols\_mallows\_cp}\NormalTok{(fit\_adjr\_without,trans\_fit) }\CommentTok{\#{-}14.49}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -14.49878
\end{verbatim}

\hypertarget{aic}{%
\paragraph{AIC}\label{aic}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{AIC}\NormalTok{(fit\_back\_without) }\CommentTok{\# 1371.373}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1371.373
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{AIC}\NormalTok{(trans\_fit\_without) }\CommentTok{\#1375.258}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1375.258
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{AIC}\NormalTok{(fit\_adjr\_without) }\CommentTok{\# 1371.363}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1371.363
\end{verbatim}

\hypertarget{bic}{%
\paragraph{BIC}\label{bic}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{BIC}\NormalTok{(fit\_back\_without) }\CommentTok{\#1424.441}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1424.441
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{BIC}\NormalTok{(trans\_fit\_without) }\CommentTok{\#1440,573}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1440.573
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{BIC}\NormalTok{(fit\_adjr\_without) }\CommentTok{\# 1428.514}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1428.514
\end{verbatim}

\hypertarget{mse}{%
\paragraph{MSE}\label{mse}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{back\_without\_summ }\OtherTok{=}\FunctionTok{summary}\NormalTok{(fit\_back\_without) }\CommentTok{\#1.2633}
\FunctionTok{mean}\NormalTok{(back\_without\_summ}\SpecialCharTok{$}\NormalTok{residuals}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.263328
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{trans\_without\_summ }\OtherTok{=}\FunctionTok{summary}\NormalTok{(trans\_fit\_without) }\CommentTok{\#1.2572}
\FunctionTok{mean}\NormalTok{(trans\_without\_summ}\SpecialCharTok{$}\NormalTok{residuals}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.257243
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{adjr\_without\_summ }\OtherTok{=}\FunctionTok{summary}\NormalTok{(fit\_adjr\_without) }\CommentTok{\#1.2575}
\FunctionTok{mean}\NormalTok{(adjr\_without\_summ}\SpecialCharTok{$}\NormalTok{residuals}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.257546
\end{verbatim}

\hypertarget{model-validation}{%
\section{Model Validation}\label{model-validation}}

\hypertarget{compute-rmse-mspe-adjusted-r2-by-cross-validation}{%
\subsection{Compute RMSE, MSPE, adjusted R\^{}2 by
cross-validation}\label{compute-rmse-mspe-adjusted-r2-by-cross-validation}}

Using back and adjr model and data without outlier

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(modelr)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{cv\_df }\OtherTok{=} 
  \FunctionTok{crossv\_kfold}\NormalTok{(cdi\_trans\_without, }\AttributeTok{k =} \DecValTok{5}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{train =} \FunctionTok{map}\NormalTok{(train, as\_tibble),}
    \AttributeTok{test =} \FunctionTok{map}\NormalTok{(test, as\_tibble))}

\NormalTok{cv\_df }\OtherTok{=} 
\NormalTok{  cv\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{fit\_back =} \FunctionTok{map}\NormalTok{(train, }\SpecialCharTok{\textasciitilde{}}\FunctionTok{lm}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{ pop }\SpecialCharTok{+}\NormalTok{ pop18 }\SpecialCharTok{+}\NormalTok{ hsgrad }\SpecialCharTok{+}\NormalTok{ bagrad }\SpecialCharTok{+}\NormalTok{ poverty }\SpecialCharTok{+}\NormalTok{ pcincome }\SpecialCharTok{+}\NormalTok{ region }\SpecialCharTok{+}\NormalTok{ pbeds\_1000 }\SpecialCharTok{+}\NormalTok{ density\_pop, }\AttributeTok{data =}\NormalTok{ .x)),}
    \AttributeTok{trans\_fit =} \FunctionTok{map}\NormalTok{(train, }\SpecialCharTok{\textasciitilde{}}\FunctionTok{lm}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{ pop }\SpecialCharTok{+}\NormalTok{ pop18 }\SpecialCharTok{+}\NormalTok{ pop65}\SpecialCharTok{+}\NormalTok{ hsgrad }\SpecialCharTok{+}\NormalTok{ bagrad }\SpecialCharTok{+}\NormalTok{ poverty }\SpecialCharTok{+}\NormalTok{ unemp }\SpecialCharTok{+}\NormalTok{ pcincome }\SpecialCharTok{+}\NormalTok{ region}\SpecialCharTok{+} \SpecialCharTok{+}\NormalTok{pdocs\_1000}\SpecialCharTok{+}\NormalTok{pbeds\_1000 }\SpecialCharTok{+}\NormalTok{ density\_pop, }\AttributeTok{data =}\NormalTok{ .x)),}
    \AttributeTok{fit\_adjr =} \FunctionTok{map}\NormalTok{(train, }\SpecialCharTok{\textasciitilde{}}\FunctionTok{lm}\NormalTok{(crm\_1000\_sqr }\SpecialCharTok{\textasciitilde{}}\NormalTok{ pop }\SpecialCharTok{+}\NormalTok{ pop18 }\SpecialCharTok{+}\NormalTok{ hsgrad }\SpecialCharTok{+}\NormalTok{ bagrad }\SpecialCharTok{+}\NormalTok{ poverty }\SpecialCharTok{+}\NormalTok{ unemp }\SpecialCharTok{+}\NormalTok{ pcincome }\SpecialCharTok{+}\NormalTok{ region}\SpecialCharTok{+}\NormalTok{ pbeds\_1000 }\SpecialCharTok{+}\NormalTok{ density\_pop, }\AttributeTok{data =}\NormalTok{ .x))) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{rmse\_back =} \FunctionTok{map2\_dbl}\NormalTok{(fit\_back, test, }\SpecialCharTok{\textasciitilde{}}\FunctionTok{rmse}\NormalTok{(}\AttributeTok{model =}\NormalTok{ .x, }\AttributeTok{data =}\NormalTok{ .y)),}
    \AttributeTok{rmse\_full =} \FunctionTok{map2\_dbl}\NormalTok{(trans\_fit, test, }\SpecialCharTok{\textasciitilde{}}\FunctionTok{rmse}\NormalTok{(}\AttributeTok{model =}\NormalTok{ .x, }\AttributeTok{data =}\NormalTok{ .y)),}
    \AttributeTok{rmse\_adjr =} \FunctionTok{map2\_dbl}\NormalTok{(fit\_adjr, test, }\SpecialCharTok{\textasciitilde{}}\FunctionTok{rmse}\NormalTok{(}\AttributeTok{model =}\NormalTok{ .x, }\AttributeTok{data =}\NormalTok{ .y))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{mspe\_back =} \FunctionTok{map2\_dbl}\NormalTok{(fit\_back, test, }\SpecialCharTok{\textasciitilde{}} \FunctionTok{mean}\NormalTok{((.y}\SpecialCharTok{$}\NormalTok{crm\_1000\_sqr }\SpecialCharTok{{-}}\FunctionTok{predict.lm}\NormalTok{(.x,.y))}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)),}
    \AttributeTok{mspe\_full =} \FunctionTok{map2\_dbl}\NormalTok{(trans\_fit, test, }\SpecialCharTok{\textasciitilde{}} \FunctionTok{mean}\NormalTok{((.y}\SpecialCharTok{$}\NormalTok{crm\_1000\_sqr }\SpecialCharTok{{-}}\FunctionTok{predict.lm}\NormalTok{(.x,.y))}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)),}
    \AttributeTok{mspe\_adjr =} \FunctionTok{map2\_dbl}\NormalTok{(fit\_adjr, test, }\SpecialCharTok{\textasciitilde{}} \FunctionTok{mean}\NormalTok{((.y}\SpecialCharTok{$}\NormalTok{crm\_1000\_sqr }\SpecialCharTok{{-}}\FunctionTok{predict.lm}\NormalTok{(.x,.y))}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{res\_back =} \FunctionTok{map}\NormalTok{(fit\_back, broom}\SpecialCharTok{::}\NormalTok{glance }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ as.data.frame),}
    \AttributeTok{res\_full =} \FunctionTok{map}\NormalTok{(trans\_fit, broom}\SpecialCharTok{::}\NormalTok{glance }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ as.data.frame),}
    \AttributeTok{res\_adjr =} \FunctionTok{map}\NormalTok{(fit\_adjr, broom}\SpecialCharTok{::}\NormalTok{glance }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ as.data.frame)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{unnest}\NormalTok{(res\_back, res\_full, res\_adjr) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(rmse\_back,rmse\_full,rmse\_adjr,}
\NormalTok{                mspe\_back,mspe\_full,mspe\_adjr,}
\NormalTok{                value.adj.r.squared,value.adj.r.squared1,value.adj.r.squared2) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rename}\NormalTok{(}\AttributeTok{adjR\_back =}\NormalTok{ value.adj.r.squared,}
         \AttributeTok{adjR\_full =}\NormalTok{ value.adj.r.squared1,}
         \AttributeTok{adjR\_adjr =}\NormalTok{ value.adj.r.squared2)}

\NormalTok{cv\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarise\_each}\NormalTok{(}\FunctionTok{funs}\NormalTok{(}\FunctionTok{mean}\NormalTok{( .,}\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{t}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                [,1]
## rmse_back 1.1675477
## rmse_full 1.1812447
## rmse_adjr 1.1656535
## mspe_back 1.3695900
## mspe_full 1.4017130
## mspe_adjr 1.3651876
## adjR_back 0.5378246
## adjR_full 0.5371541
## adjR_adjr 0.5386250
\end{verbatim}

\hypertarget{plot-the-violin-plot}{%
\subsection{Plot the violin plot}\label{plot-the-violin-plot}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{unnest\_cd\_df }\OtherTok{=}\NormalTok{ cv\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(rmse\_back}\SpecialCharTok{:}\NormalTok{adjR\_adjr, }
               \AttributeTok{names\_pattern =} \StringTok{"(.*)(....)$"}\NormalTok{, }
               \AttributeTok{names\_to =} \FunctionTok{c}\NormalTok{(}\StringTok{"limit"}\NormalTok{, }\StringTok{"model"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{limit=}\FunctionTok{ifelse}\NormalTok{(limit}\SpecialCharTok{==}\StringTok{""}\NormalTok{, }\StringTok{"value"}\NormalTok{, limit)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{id\_cols =}\NormalTok{ model, }
              \AttributeTok{names\_from =}\NormalTok{ limit, }
              \AttributeTok{values\_from =}\NormalTok{ value, }
              \AttributeTok{names\_repair =} \StringTok{"check\_unique"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{unnest}\NormalTok{(}\FunctionTok{c}\NormalTok{(rmse\_, adjR\_)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{rename}\NormalTok{(}\AttributeTok{rmse =}\NormalTok{ rmse\_,}
         \AttributeTok{adjR =}\NormalTok{ adjR\_)}
\NormalTok{unnest\_cd\_df }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ model, }\AttributeTok{y =}\NormalTok{ rmse)) }\SpecialCharTok{+} \FunctionTok{geom\_violin}\NormalTok{() }
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-43-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{unnest\_cd\_df }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ model, }\AttributeTok{y =}\NormalTok{ adjR)) }\SpecialCharTok{+} \FunctionTok{geom\_violin}\NormalTok{() }
\end{Highlighting}
\end{Shaded}

\includegraphics{final_project_files/figure-latex/unnamed-chunk-43-2.pdf}

\end{document}

### 2021-12-17 final_project.Rmd
correct cp

@@
 ```{r}
-olsrr::ols_mallows_cp(fit_back_without,trans_fit) #-14.65
-olsrr::ols_mallows_cp(trans_fit_without,trans_fit)# -10.59
-olsrr::ols_mallows_cp(fit_adjr_without,trans_fit) #-14.49
+olsrr::ols_mallows_cp(fit_back_without,trans_fit_without) # 7.047
+olsrr::ols_mallows_cp(trans_fit_without,trans_fit_without)# 11
+olsrr::ols_mallows_cp(fit_adjr_without,trans_fit_without) # 7.102
 ```

### 2021-12-17 draft_yma.Rmd
delete stuff i don't need

@@
 library(readr)
+library(flexmix)
 ```
@@
   data = cdi_descriptive,
-  type = "parametric", # parametric for Pearson, nonparametric for Spearman's correlation
+  type = "parametric", 
   colors = c("darkred", "white", "steelblue") # change default colors
 )
+
+ggcorrplot(cor(cdi_descriptive), type = "upper",
+   lab = TRUE)
 ```
@@
 
-### New tibble with transformed crm_1000 -> crm_1000_sqr
+### New tibble with transformed y
 ```{r}
@@
 # cdi %>% filter(pop == 2300664 | pop == 396685) %>% select(cty,pop)
+
+cdi_trans_without1 = cdi_tans[]
 ```
@@
 
+```{r}
+fit_back_int = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop + pcincome*poverty, data = cdi_trans)
+summary(fit_back_int)
+
+olsrr::ols_plot_resid_fit(fit_back_int)
+olsrr::ols_plot_resid_qq(fit_back_int)
+MASS::boxcox(fit_back_int)
+#Cp
+olsrr::ols_mallows_cp(fit_back_int,trans_fit) #13.94
+#adjr 0.5701
+
+BIC(fit_back_int) #1435.05
+
+back_int_summ =summary(fit_back_int) #1.331
+mean(back_int_summ$residuals^2) #1.2598
+
+performance::check_collinearity(fit_back_int)
+
+fit_back_int_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop + pcincome*poverty, data = cdi_trans_without)
+
+summary(fit_back_int_without) #adjr 0.5658
+
+# Residual performance
+par(mfrow=c(2,2))
+olsrr::ols_plot_resid_fit(fit_back_int_without)
+olsrr::ols_plot_resid_qq(fit_back_int_without)
+MASS::boxcox(fit_back_int_without)
+
+#Cp -38.65??
+olsrr::ols_mallows_cp(fit_back_int_without,trans_fit)
+#BIC: 1401.441
+BIC(fit_back_int_without) 
+#MSE
+back_int_without_summ =summary(fit_back_int_without) #1.18
+mean(back_int_without_summ$residuals^2)
+#VIF 
+performance::check_collinearity(fit_back_int_without)
+```
+
 ### Criterion
@@
 summary(fit_Cp) #0.5469
+```
+
+#### AIC
+
+```{r}
+AIC(fit_back) # 1400.399
+AIC(fit_forward) #1404.519
+AIC(fit_both) #1400.399
+AIC(fit_adjr) # 1400.889
+AIC(fit_Cp) # 1400.399
 ```
@@
 
+## Checking to Outliers and Influential Points
+
 ```{r}
-par(mfrow=c(2,2))
-plot(fit_back)
+olsrr::ols_plot_cooksd_bar(strp_back, print_plot = TRUE)
+#remove observations whose cook's D > 0.5
+cdi_trans_without = cdi_trans[-c(1, 6),]
 ```
 
-## Checking to Outliers and Influential Points
-
-### Back
-
-#### Old Method
-```{r}
-# residuals vs leverage plot
-plot(fit_back, which = 4)
-#remove outliers
-cdi_trans_without = cdi_trans[-c(1, 6, 53),]
-fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + 
-    pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
-summary(fit_without)
-```
-
-```{r}
-par(mfrow=c(2,2))
-plot(fit_without)
-```
-
-### Threshold: $D_{i}>\frac{4}{n} = \frac{4}{440} = 0.009$
-```{r}
-cook = olsrr::ols_plot_cooksd_bar(fit_back, print_plot = TRUE)
-normal = cook$data$color == "normal"
-outlier = cook$data$color == "outlier"
-```
-
-```{r}
-outlier_df = cdi[outlier,]
-```
-
-This gives us 21 outliers. We should consider if we want to take them all out
-
 ### Build model without outlier
-
-Here's what happens when we take them all out.
-
-```{r}
-cdi_trans_without = cdi_trans[normal,]
-```
 
@@
 ```{r}
-fit_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
+fit_back_without = lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop, data = cdi_trans_without)
 
@@
 par(mfrow=c(2,2))
-olsrr::ols_plot_resid_fit(fit_without)
-olsrr::ols_plot_resid_qq(fit_without)
-MASS::boxcox(fit_without)
+olsrr::ols_plot_resid_fit(fit_back_without)
+olsrr::ols_plot_resid_qq(fit_back_without)
+MASS::boxcox(fit_back_without)
+```
 
-#Cp -88.45??
-olsrr::ols_mallows_cp(fit_without,trans_fit)
-#BIC: 1274.534
-BIC(fit_without) 
-#MSE
-without_summ =summary(fit_without) #1.017
-mean(without_summ$residuals^2)
-#VIF 
-performance::check_collinearity(fit_without)
+Build the without model with **trans_fit** (full model)
+
+```{r}
+trans_fit_without = lm(crm_1000_sqr ~., data = cdi_trans_without)
+
+summary(trans_fit_without)
+
+# Residual performance
+par(mfrow=c(2,2))
+olsrr::ols_plot_resid_fit(trans_fit_without)
+olsrr::ols_plot_resid_qq(trans_fit_without)
+MASS::boxcox(trans_fit_without)
 ```
@@
 par(mfrow=c(2,2))
-olsrr::ols_plot_resid_fit(fit_without2)
-olsrr::ols_plot_resid_qq(fit_without2)
-MASS::boxcox(fit_without2)
+olsrr::ols_plot_resid_fit(fit_adjr_without)
+olsrr::ols_plot_resid_qq(fit_adjr_without)
+MASS::boxcox(fit_adjr_without)
+```
 
-#Cp -87.5??
-olsrr::ols_mallows_cp(fit_without2,trans_fit)
-#BIC: 1279.154
-BIC(fit_without2)
-#MSE
-without2_summ =summary(fit_without2) #1.013
-mean(without2_summ$residuals^2)
-#VIF 
-performance::check_collinearity(fit_without2) 
+#### Cp for models fit without outlier
+
+```{r}
+olsrr::ols_mallows_cp(fit_back_without,trans_fit_without) # 7.047
+olsrr::ols_mallows_cp(trans_fit_without,trans_fit_without)# 11
+olsrr::ols_mallows_cp(fit_adjr_without,trans_fit_without) # 7.102
+```
+
+#### AIC
+
+```{r}
+AIC(fit_back_without) # 1371.373
+AIC(trans_fit_without) #1375.258
+AIC(fit_adjr_without) # 1371.363
+AIC(fit_back_int_without) #1344.376
+```
+
+#### BIC
+
+```{r}
+BIC(fit_back_without) #1424.441
+BIC(trans_fit_without) #1440,573
+BIC(fit_adjr_without) # 1428.514
+BIC(fit_back_int_without) #1401.527
+```
+
+#### MSE
+
+```{r}
+back_without_summ =summary(fit_back_without) #1.2633
+mean(back_without_summ$residuals^2)
+
+trans_without_summ =summary(trans_fit_without) #1.2572
+mean(trans_without_summ$residuals^2)
+
+adjr_without_summ =summary(fit_adjr_without) #1.2575
+mean(adjr_without_summ$residuals^2)
+
+int_without_summ =summary(fit_back_int_without) #1.18
+mean(int_without_summ$residuals^2)
+
 ```
@@
     fit_adjr = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x)),
-    fit_forward = map(train, ~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + unemp + pcincome + region+ pbeds_1000 + density_pop, data = .x))) %>% 
-  mutate(
-    rmse_model2 = map2_dbl(fit_back, test, ~rmse(model = .x, data = .y)),
-    rmse_model3 = map2_dbl(trans_fit, test, ~rmse(model = .x, data = .y)),
-    rmse_model4 = map2_dbl(fit_adjr, test, ~rmse(model = .x, data = .y)),
-    rmse_model5 = map2_dbl(fit_forward, test, ~rmse(model = .x, data = .y))) %>%
-  mutate(
-    mspe_model2 = map2_dbl(fit_back, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
-    mspe_model3 = map2_dbl(trans_fit, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
-    mspe_model4 = map2_dbl(fit_adjr, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
-    mspe_model5 = map2_dbl(fit_forward, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2))
+    fit_back_int_without = map(train,~lm(crm_1000_sqr ~ pop + pop18 + hsgrad + bagrad + poverty + pcincome + region + pbeds_1000 + density_pop + pcincome*poverty, data = .x))
     ) %>% 
   mutate(
-    res_model2 = map(fit_back, broom::glance %>% as.data.frame),
-    res_model3 = map(trans_fit, broom::glance %>% as.data.frame),
-    res_model4 = map(fit_adjr, broom::glance %>% as.data.frame),
-    res_model5 = map(fit_forward, broom::glance %>% as.data.frame)) %>%
-  unnest(res_model2, res_model3, res_model4, res_model5) %>%
-  dplyr::select(rmse_model2,rmse_model3,rmse_model4,rmse_model5,
-                mspe_model2,mspe_model3,mspe_model4,mspe_model5,
+    rmse_back = map2_dbl(fit_back, test, ~rmse(model = .x, data = .y)),
+    rmse_full = map2_dbl(trans_fit, test, ~rmse(model = .x, data = .y)),
+    rmse_adjr = map2_dbl(fit_adjr, test, ~rmse(model = .x, data = .y)),
+    rmse_bain = map2_dbl(fit_back_int_without, test, ~rmse(model = .x, data = .y))) %>%
+  mutate(
+    mspe_back = map2_dbl(fit_back, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
+    mspe_full = map2_dbl(trans_fit, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
+    mspe_adjr = map2_dbl(fit_adjr, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2)),
+    mspe_bain = map2_dbl(fit_back_int_without, test, ~ mean((.y$crm_1000_sqr -predict.lm(.x,.y))^2))
+  ) %>% 
+  mutate(
+    res_back = map(fit_back, broom::glance %>% as.data.frame),
+    res_full = map(trans_fit, broom::glance %>% as.data.frame),
+    res_adjr = map(fit_adjr, broom::glance %>% as.data.frame),
+    res_bain = map(fit_back_int_without, broom::glance %>% as.data.frame),
+    ) %>%
+  unnest(res_back, res_full, res_adjr, res_bain) %>%
+  dplyr::select(rmse_back,rmse_full,rmse_adjr,rmse_bain,
+                mspe_back,mspe_full,mspe_adjr,mspe_bain,
                 value.adj.r.squared,value.adj.r.squared1,value.adj.r.squared2,value.adj.r.squared3) %>%
-  rename(adjR_model2 = value.adj.r.squared,
-         adjR_model3 = value.adj.r.squared1,
-         adjR_model4 = value.adj.r.squared2,
-         adjR_model5 = value.adj.r.squared3)
+  rename(adjR_back = value.adj.r.squared,
+         adjR_full = value.adj.r.squared1,
+         adjR_adjr = value.adj.r.squared2,
+         adjR_bain = value.adj.r.squared3)
 
@@
 unnest_cd_df = cv_df %>%
-  pivot_longer(rmse_model2:adjR_model5, 
-               names_pattern = "(.*)(......)$", 
+  pivot_longer(rmse_back:adjR_bain, 
+               names_pattern = "(.*)(....)$", 
                names_to = c("limit", "model")) %>% 
@@
 
-## backward by hand
+```{r}
+olsrr::ols_plot_resid_qq(fit_back_int_without)
+olsrr::ols_plot_resid_fit(fit_back_int_without)
+MASS::boxcox(fit_back_int_without)
+```
 
 ```{r}
-# No docs
-step1 = update(trans_fit, . ~ . -pdocs_1000)
-summary(step1)
+locfit::cp(fit_back_without)
+```
 
-# No pop65
-step2 = update(step1, . ~ . -pop65)
-summary(step2)
+```{r}
+models1 <- regsubsets(crm_1000_sqr ~., data = cdi_trans_without,nvmax = 12)
+res_sum1 = summary(models1)
 
-# No unemp
-step3 = update(step2, . ~ . -unemp)
-summary(step3)
-
-step4 = update(step3, . ~ . -hsgrad)
-summary(step4)
-
-mult.fit.final = lm(crm_1000^0.5 ~ pop + pop18 + bagrad + poverty + 
-    pcincome + region + pbeds_1000 + density_pop, data = cdi_model)
-summary(mult.fit.final)
+par(mfrow=c(1,2))
+plot(1:12, res_sum1$cp, xlab="No of parameters", ylab="Cp Statistic")
+plot(1:12, res_sum1$adjr2, xlab="No of parameters", ylab="Adj R2")
 ```
